## Empirical Evidence

```{r isotopesAndTemperature, echo=FALSE}
question("Which statement is correct?",
  answer("Larger relative abundances of $^{18}$O in the air correspond to warmer global temperatures"),
  answer("Larger relative abundances of $^{16}$O in the air correspond to warmer global temperatures"),
  answer("Larger relative abundances of $^{18}$O in the air correspond to colder global temperatures", correct = TRUE)
)
```

```{r sortEra, echo = FALSE}

era <- c("Cenozoic", "Mesozoic", "Paleozoic")
rank_era <- sortable::question_rank(
  "Sort the three eras by age starting with the youngest era first",
  learnr::answer(era, correct = TRUE),
  learnr::answer(rev(era), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_era
```

```{r icehouseVsGreenhouse, echo=FALSE}
question("Which statement is correct?",
  answer("We currently live in a greenhouse period"),
  answer("We currently live in an icehouse period", correct = TRUE)
)
```


```{r summary, echo=FALSE}
question("Which statement is correct? Select all that apply",
  answer("Current global mean temperatures exceed temperatures of the Cenozoic era"),
  answer("The earth has been both, much warmer and much colder than today", correct = TRUE),
  answer("Temperatures in the holocene have been declining for about 7,000 years until about the 1800", correct = TRUE)
)
```

### Exercise (1)

* The government has invited you to give a brief presentation assessing whether the current and projected future atmospheric CO~2~ concentrations have historical precedents in Earth's past. Prepare your presentation by following the steps below.

* Download paleoclimate data of atmospheric CO~2~ concentration from [here](https://www.annualreviews.org/doi/suppl/10.1146/annurev-earth-082420-063026) (Download Supplemental Data 1 (XLSX))
* Open ```ea49_rae_suppl_data1.xlsx``` and go to sheet ```d11B_pH```
* Save the sheet as a ```.csv``` file
* Read in the file in R using the ```read.csv()``` command:

```
data <- read.csv("C:/Users/whatever/your/path/is/ea49_rae_suppl_data1.csv", header = TRUE)
```
* The data column ```xco2``` gives the atmospheric CO2 concentration in ppm 
* Make a scatterplot of [CO~2~] and Age:

```{r ageVsCO2, exercise = TRUE, exercise.eval = FALSE}

```

```{r ageVsCO2-solution}

myFile <- file.path("data", "ea49_rae_suppl_data1.csv") 
data <- read.csv(myFile, header = TRUE)
age <- data$age / 1000 # Age in million years
co2 <- data$xco2
plot(age, co2, xlab = "Age (Ma)", ylab = "[CO2] (ppm)", xlim=rev(c(0, 70)))

```

* How does the CO~2~ variation compare with the variation in global mean temperature plotted in Figure 2.11 on page 31?

* Estimate how high CO~2~ levels were during the early Eocene climate optimum

* Go to Figure 8.3 on page 138 and estimate the CO~2~ for the year 2100 under the fossil fuel-intensive Shared Socioeconomic Pathway SSP5-8.5. Add a horizontal line that shows the projected CO~2~ concentration ``` abline(h = xxx, col = "blue")```.

* The data that you are analyzing has been published by [Rae et al (2022)](https://www.annualreviews.org/doi/pdf/10.1146/annurev-earth-082420-063026). The abstract states how far back in time one must go in order to find CO~2~ concentrations that correspond to today's level. Read the abstract and verify their statement by adding a horizontal line that shows today's CO~2~ levels.

```{r addCurrentCO2, exercise = TRUE, exercise.eval = FALSE}

```

```{r  addCurrentCO2-solution}

myFile <- file.path("data", "ea49_rae_suppl_data1.csv") 
data <- read.csv(myFile, header = TRUE)

age <- data$age / 1000
co2 <- data$xco2
plot(age, co2, xlab = "Age (Ma)", ylab = "[CO2] (ppm)", xlim=rev(c(0, 3.5)))

abline(h = 421, col = "red")

```

* Improve the readability of your plot by adjusting the time window. This can be done by changing the ```xlim=rev(c(0, 70))``` to e.g. ```xlim=rev(c(0, 10))```. Also, adjust the y-axis by adding ```ylim = c(100,500)```.


```{r  addCurrentCO2v2, exercise = TRUE, exercise.eval = FALSE}

```

```{r  addCurrentCO2v2-solution}
myFile <- file.path("data", "ea49_rae_suppl_data1.csv") 
data <- read.csv(myFile, header = TRUE)

age <- data$age / 1000
co2 <- data$xco2
plot(age, co2, xlab = "Age (Ma)", ylab = "[CO2] (ppm)", xlim=rev(c(0, 3.5)), ylim = c(100,500))

abline(h = 421, col = "red")

```

### Exercise (2)

* You are a group of scientists who just returned from Antarctica where you have conducted ice core drilling to reconstruct temperature changes during earth's history. You have been invited to share your results at the American Geophysical Union Fall Meeting. To prepare your presentation, follow the steps below.

* Some background information: The original Vostock ice core data are provided by [Petit et al (2000)](https://data.ess-dive.lbl.gov/view/doi%3A10.3334%2FCDIAC%2FCLI.006) (vostok.1999.temp.dat). The data are described here: [Petit, J.R., Jouzel, J., Raynaud, D., Barkov, N.I., Barnola, J.M., Basile, I., Bender, M., Chappellaz, J., Davis, M., Delaygue, G. and Delmotte, M., 1999. Climate and atmospheric history of the past 420,000 years from the Vostok ice core, Antarctica. Nature, 399(6735), pp.429-436.](https://www.nature.com/articles/20859). I have prepared the data for you, so ther is no need for downloading data.

* Plot ice core depth versus age:

```{r vostock1999temp, exercise = TRUE, exercise.eval = FALSE}

myFile <- file.path("data", "vostok.1999.temp.dat") 
data <- read.table(myFile, header = TRUE)

plot(x = ___,  y = ___/1000,
     xlab = "Depth (m)",
     ylab = "Age (thousands of years)",
     type = "l")

```

```{r vostock1999temp-solution}

myFile <- file.path("data", "vostok.1999.temp.dat") 
data <- read.table(myFile, header = TRUE)

plot(data$Depth,  data$Age/1000,
     xlab = "Depth (m)",
     ylab = "Age (thousands of years)",
     type = "l")
```

* How deep did you have to drill to obtain air bubbles that were trapped about 400,000 years ago?

* Next, plot the relation between Deuterium and temperature.

```{r Deuterium, exercise = TRUE, exercise.eval = FALSE}

```

```{r Deuterium-solution}
myFile <- file.path("data", "vostok.1999.temp.dat") 
data <- read.table(myFile, header = TRUE)

plot(data$Deuterium,  data$Temperature, 
     xlab = "Deuterium",
     ylab = "Temperature Anomlies (C)")
```

* Deuterium is expressed as $\delta D$ in ‰ with respect to Standard Mean Ocean Water (SMOW): $\delta D_{ice} = \left( \frac{(D/H)_{sample}}{(D/H)_{SMOW}}-1 \right) \times 1,000$, where $H$ refers to Protium.
 
* Explain what Deuterium is, what the values show, and why Deuterium and temperature are positively correlated.

* Next, plot how the temperature varies in time. To sort your plot from 400,000 to present, use ```xlim = rev(c(0, 400000))```.

```{r vostockTemp, exercise = TRUE, exercise.eval = FALSE}

```

```{r vostockTemp-solution}

myFile <- file.path("data", "vostok.1999.temp.dat") 
data <- read.table(myFile, header = TRUE)

plot(data$Age,  data$Temperature, 
     xlab = "Age (thousands of years)",
     ylab = "Temperature Anomlies (C)",
     type = "l", xlim = rev(c(0, 400000)))
```

* How many interglacial periods do you see? 
* What periods are longer, glacial or interglacial? 
* What happens faster, warming or cooling?
* Below I am showing the same data as above, but I added a second line that shows [CO~2~] in red:

```{r}
rm(list = ls())

myFile <- file.path("data", "vostok.1999.temp.dat") 
temp <- read.table(myFile, header = TRUE)

myFile <- file.path("data", "co2nat.txt") 
co2 <- read.table(myFile, header = TRUE)

# Create a sequence of years from the minimum to maximum year in your data
all_years <- seq(min(temp$Age), max(temp$Age), by = 1)

# Merge the original data with the complete sequence of years
temp_allYears <- merge(data.frame(Age = all_years), temp, all.x = TRUE)

co2_allYears <- merge(data.frame(Age = all_years), co2, all.x = TRUE)

data <- merge(temp_allYears, co2_allYears, by = "Age")

x <- data$Age / 1000
y1 <- data$Temperature
y2 <- data$CO2

my.xlim <- rev(c(0, 400))

# Create a layout with two side-by-side plots and distinct vertical axes
par(mfrow = c(1, 1), mar = c(5, 4, 4, 5))

# Plot the first variable with the left y-axis
plot(x, y1, ylab = "Temperature anomaly (C)", xlab = "Age (thousands of years)", col = NA, xlim = my.xlim)
lines(x[!is.na(y1)], y1[!is.na(y1)], col = "grey", xlim=rev(c(300, 350)), lwd = 2)

# Add a second y-axis on the right for the second variable
par(new = TRUE)
plot(x, y2, yaxt = "n", xaxt = "n", xlab = NA, ylab = NA, col = NA, xlim = my.xlim)

lines(x[!is.na(y2)], y2[!is.na(y2)], col = "red", xlim=rev(c(300, 350)), lwd = 2)


axis(4, col = "red", col.axis = "red")

# Add axis labels
mtext("[CO2] (ppmv)", side = 4, line = 3, col = "red")

```

* Comparing both lines, does temperature follow CO~2~, or does CO~2~ follow temperature?
* Try to explain your answer using the concepts of forcing and feedback.
* Considering anthropogenic climate change, is CO~2~ a forcing or a feedback?

### Exercise (3)
* Let's plot the [CO~2~] paleoclimate data from the last 400,000 years and the recent [CO~2~] observations from the Mauna Loa Observatory together.

```{r pastAndPresentCO2, exercise = TRUE, exercise.eval = FALSE}

myFile <- file.path("data", "pastAndPresentCO2.txt") 
data <- read.table(myFile, header = TRUE)

plot(x = ___/1000, y = ___, xlab = "Age (thousands of years)", ylab = "[CO2] (ppmv)")

```

```{r pastAndPresentCO2-solution}

myFile <- file.path("data", "pastAndPresentCO2.txt") 
data <- read.table(myFile, header = TRUE)

plot(data$Age/1000, data$CO2, xlab = "Age (thousands of years)", ylab = "[CO2] (ppmv)")

```


* How does the current [CO~2~] trend compare to trends in the Earth's past?
