## Radiation and Energy Balance

The energy of a photon ($E$) is a function of its wavelength ($\lambda$): $$E = h \frac{c}{\lambda}$$ where $h$ is Planck's constant ($h = 6.63 \times 10^{-34}$ J s) and $c$ is the speed of light ($c = 3 \times 10^8$ m s$^{-1}$). Plot the energy of a photon as a function of wavelength, ranging from 0.1 to 2.0 $\mu$m.

```{r photon, exercise = TRUE, exercise.eval = FALSE}

# constants
h <- ___ # Planck's constant (J s)
c <- ___ # speed of light (m s-1)

# variable
lambda <- ___ # wavelength (m)

# function
E <- ___

# plot
my.xlab <- "Wavelength ($mu$m)"
my.ylab <- "Energy of a photon (J)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = lambda, y = E, type = "l", lty = 1, xlab = my.xlab, ylab = my.ylab)

```

```{r photon-solution}
# constants
h <- 6.63 * 10^(-34) # Planck's constant (J s)
c <- 3 * 10^8 # speed of light (m s-1)

# variable
lambda <- seq(0.1,2,0.01) * 10^(-6) # wavelength (m)

# function
E <- h * c / lambda

# plot
my.xlab <- "Wavelength ($mu$m)"
my.ylab <- "Energy of a photon (J)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = lambda, y = E, type = "l", lty = 1, xlab = my.xlab, ylab = my.ylab)

```

Planck's function of spectral irradiance is given as:

$$ I(\lambda) = \frac{2 \pi h c^2}{\lambda^5}\frac{1}{\exp\left(  \frac{h c}{\lambda k_B T} \right)-1}$$ Plot this function for wavelengths ranging from 0 to 1 $\mu$m and for different temperatures between 3000 and 6000 Kelvin.

```{r PlanckFunction, exercise = TRUE, exercise.eval = FALSE}
# constants
h <- ___ # Planck's constant (J s)
kB <- ___ # Boltzmann constant (J K-1)
c <- ___ # speed of light (m s-1)

# variables
T <- ___ # temperature (K)
lambda <- ___ # wavelength (m)

# Planck's function

my.fun <- function(lambda, T){
  I <- ___
  return(I)
}

data <- outer(lambda, T, my.fun)

my.xlab <- "Wavelength ($mu$m)"
my.ylab <- "Spectral Irradiance (W m$^{-2} mu$m$^{-1}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- rev(viridis::plasma(n = ncol(data)))

par(mar = c(5,5,1,1))
matplot(x = lambda * 10^6, y = data / 10^6, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("T =", T, "K", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r PlanckFunction-solution}
# constants
h <- 6.63 * 10^(-34) # Planck's constant (J s)
kB <- 1.38 * 10^(-23) # Boltzmann constant (J K-1)
c <- 3 * 10^8 # speed of light (m s-1)

# variables
T <- seq(3000,6000, 500) # temperature (K)
lambda <- seq(0,2,0.01) * 10^(-6) # wavelength (m)

# Planck's function

my.fun <- function(lambda, T){
  termI <- 2 * pi * h * c^2 / lambda^5
  termII <- exp(h * c / (lambda * kB * T)) - 1
  I <- termI / termII
  return(I)
}

data <- outer(lambda, T, my.fun)

my.xlab <- "Wavelength ($mu$m)"
my.ylab <- "Spectral Irradiance (W m$^{-2} mu$m$^{-1}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- rev(viridis::plasma(n = ncol(data)))

par(mar = c(5,5,1,1))
matplot(x = lambda * 10^6, y = data / 10^6, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("T =", T, "K", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")
```

Plot Wien's law for a temperature range of 100 to 6000 Kelvin.

```{r WiensLaw, exercise = TRUE, exercise.eval = FALSE}

T <- ___
lambdaMax <- ___

my.xlab <- "Temperature (K)"
my.ylab <- "Wavelength ($mu$m)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = T, y = lambdaMax, type = "l", lty = 1, xlab = my.xlab, ylab = my.ylab)

```

```{r WiensLaw-solution}

T <- seq(100,6000,1)
lambdaMax <- 2897 / T

my.xlab <- "Temperature (K)"
my.ylab <- "Wavelength ($mu$m)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = T, y = lambdaMax, type = "l", lty = 1, xlab = my.xlab, ylab = my.ylab)
```

Using the Stefan-Boltzmann law, show how the power per area emitted by a body as a function of temperature, ranging from 100K to 6000K.

```{r StefanBoltzmannLaw, exercise = TRUE, exercise.eval = FALSE}

# constant
sigma <- ___ # Stefan-Boltzmann constant (W m-2 K-4)

# variable
T <- seq(100,6000,1)

# function

E <- ___

my.xlab <- "Temperature (K)"
my.ylab <- "Irradiance (W m$^{-2}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = T, y = E, type = "l", lty = 1, xlab = my.xlab, ylab = my.ylab)

```

```{r StefanBoltzmannLaw-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)

# variable
T <- seq(100,6000,1)

# function

E <- sigma * T^4

my.xlab <- "Temperature (K)"
my.ylab <- "Irradiance (W m$^{-2}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = T, y = E, type = "l", lty = 1, xlab = my.xlab, ylab = my.ylab)
```

A sphere with a radius of 1m has temperature of 100$^{\circ}$C. How much power is it radiating?

```{r powerSphere, exercise = TRUE, exercise.eval = FALSE}

```

```{r powerSphere-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)
radius <- 1 # radius of sphere (m)

# variable
T <- 100 + 273.15 # temperature (K)
A <- 4 * pi * radius^2


E <- sigma * T^4 # W m-2

P <- A * E

print(P)
```

```{r powerSphere-check}
grade_result(
  pass_if(~identical(round(.result), round(13814.23)))
)
```

Consider two stars that have the spectra shown below.

```{r twoStars}
# constants
h <- 6.63 * 10^(-34) # Planck's constant (J s)
kB <- 1.38 * 10^(-23) # Boltzmann constant (J K-1)
c <- 3 * 10^8 # speed of light (m s-1)

# variables
T <- c(5000, 7000) # temperature (K)
lambda <- seq(0,3,0.01) * 10^(-6) # wavelength (m)

# Planck's function

my.fun <- function(lambda, T){
  termI <- 2 * pi * h * c^2 / lambda^5
  termII <- exp(h * c / (lambda * kB * T)) - 1
  I <- termI / termII
  return(I)
}

data <- outer(lambda, T, my.fun)

my.xlab <- "Wavelength ($mu$m)"
my.ylab <- "Spectral Irradiance (W m$^{-2} mu$m$^{-1}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- c("black", "red")

par(mar = c(5,5,1,1))
matplot(x = lambda * 10^6, y = data / 10^6, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

lambdaMax1 <- 2897 / T[1]
lambdaMax2 <- 2897 / T[2]

abline(v = lambdaMax1, col = my.col[1], lty = 2)
abline(v = lambdaMax2, col = my.col[2], lty = 2)

# legend.text <- paste("T =", T, "K", sep = " ")
# legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")
```

```{r starTemperature01, echo=FALSE}
question_radio(
  "What is the temperature of the star with the spectrum given by the red line?",
  answer("5000K"),
  answer("6000K"),
  answer("7000K", correct = TRUE),
  answer("8000K"),
  allow_retry = TRUE
)
# lambda_max = 2897 m K / T (p. 42)
```

```{r tstarTemperature02, echo=FALSE}
question_radio(
  "What is the temperature of the star with the spectrum given by the black line?",
  answer("5000K", correct = TRUE),
  answer("6000K"),
  answer("7000K"),
  answer("8000K"),
  allow_retry = TRUE
)
# lambda_max = 2897 m K / T (p. 42)
```

The surface of our sun has a mean temperature of about 5778K. The sun's radius is about 695,700 km. How much power is our sun radiating?

```{r powerSun, exercise = TRUE, exercise.eval = FALSE}

```

```{r powerSun-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)
radius <- 6.957 * 10^8 # 7 * 10^8 # radius of sun (m)

# variables
T <- 5778 # temperature (K)
A <- 4 * pi * radius^2

# Stefan-Boltzmann Law
E <- sigma * T^4 # (W m-2)

# Total power
P <- A * E # (W)

print(P)
```

```{r powerSun-check}
grade_result(
  pass_if(~identical(round(.result/10^25), round(3.843679e+26/10^25)))
)
```

The surface temperature of the sun is approximately 6000K. What is the characteristic wavelength of the sun expressed in micro meters?

```{r wavelengthSun, exercise = TRUE, exercise.eval = FALSE}

```

```{r wavelengthSun-solution}
# variable
T <- 5778 # temperature (K)
lambda <- 2.897 * 10^(-3) / T # (m)
lambda <- lambda * 10^6 # (micro meters)
print(lambda)
```

```{r wavelengthSun-check}
grade_result(
  pass_if(~identical(round(.result,1), round(0.5013846,1)))
)
```

Imagine that the sun had a radius twice as large as it does at present, but the sun emitted the same total amount of energy. What temperature would the sun have to be?

```{r temperatureSun, exercise = TRUE, exercise.eval = FALSE}

```

```{r temperatureSun-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)
radius <- 6.957 * 10^8 # radius of sun (m)

# variable
T1 <- 5778 # temperature (K)
A <- 4 * pi * radius^2
E <- sigma * T1^4 # W m-2
P <- A * E # this is the power with the actual sun radius

# variable
denominator <- sigma * 4 * pi * (2 * radius)^2
T <- (P / denominator)^(1/4)
print(T)
```

```{r temperatureSun-check}
grade_result(
  pass_if(~identical(round(.result), round(4085.663)))
)
```

Calculate the solar constant. Assume that the sun's radius equals 6.957 $\times$ 10$^8$ m and the temperature of the photosphere equals 5778K.

```{r solarConstant, exercise = TRUE, exercise.eval = FALSE}

```

```{r solarConstant-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)
radius <- 6.957 * 10^8 # radius of sun (m)

# variables
T <- 5778 # temperature (K)
A <- 4 * pi * radius^2

# Stefan-Boltzmann Law
E <- sigma * T^4 # (W m-2)

# Total power
P <- A * E # (W)

# Earth-sun distance
earth_sun_distance <- 1.496 * 10^11 # (m)

A2 <- 4 * pi * earth_sun_distance^2

S <- P / A2 # W m-2

print(S)
```

```{r solarConstant-check}
grade_result(
  pass_if(~identical(round(.result), round(1366.703)))
)
```

Estimate the earth's surface temperature using the n-layer model and the following inputs: solar constant: 1360 W m^-2^, planetary albedo = 0.3 (-), n = 0.65.

```{r nLayer, exercise = TRUE, exercise.eval = FALSE}

```

```{r nLayer-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)
S <- 1360 # Solar constant (W m-2)
alpha <- 0.3 # planetary albedo (-)
n <- 0.65

# Equation 4.5, p. 62
num <- (n + 1) * S * (1 - alpha)
denom <- 4 * sigma

T <- (num / denom)^0.25 #surface temperature (K)

print(T)
```

```{r nLayer-check}
grade_result(
  pass_if(~identical(round(.result), round(288.4825)))
)
```

Using the same data as in the exercise above, calculate the impact of the natural greenhouse gas effect on the earth's surface temperature.

```{r naturalGHG, exercise = TRUE, exercise.eval = FALSE}

```

```{r naturalGHG-solution}
# constant
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)
S <- 1360 # Solar constant (W m-2)
alpha <- 0.3 # planetary albedo (-)

# temperature without GHG (Equation 4.5, p. 62)
n <- 0
num <- (n + 1) * S * (1 - alpha)
denom <- 4 * sigma
T1 <- (num / denom)^0.25 #surface temperature (K)


# temperature with GHG (Equation 4.5, p. 62)
n <- 0.65
num <- (n + 1) * S * (1 - alpha)
denom <- 4 * sigma
T2 <- (num / denom)^0.25 #surface temperature (K)

# Difference
deltaT <- T2 - T1

print(deltaT)
```

```{r naturalGHG-check}
grade_result(
  pass_if(~identical(round(.result), round(33.94693)))
)
```

