## The Carbon Cycle
### Exercise (1)
* Make a 2-box model that simulates the exchange of carbon between the atmosphere and the biosphere. For the initial time step, put all the carbon into the atmosphere and run the model until it reaches equilibrium. Build your model based on Figure 1 and Table 1 provided by [Avenhaus, R., S. Fenyi, and H. Frick. 1979. “Box Models for the CO2 Cycle of the Earth.” Environment International 2 (4): 379–85.](https://www.sciencedirect.com/science/article/abs/pii/0160412079900114).

```{r 2boxModel, exercise = TRUE, exercise.eval = FALSE}

# Reference: 
# Avenhaus, R., S. Fenyi, and H. Frick. 1979. “Box Models for the CO2 Cycle of the Earth.” Environment International 2 (4): 379–85.

library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Parameters
    k_ab <- ___  # Transfer rate from atmosphere to biosphere measured in yr-1 (Avenhaus et al., 1979; Table 1)
    k_ba <- ___ # Transfer rate from biosphere to atmosphere measured in yr-1 (Avenhaus et al., 1979; Table 1)
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- ___
    dB_dt <- ___

    # Return the derivatives
    return(list(c(dA_dt, dB_dt)))
  })
}

# Initial state (Avenhaus et al., 1979; Table 1)
Af <- 1 # 
Bf <- 1.2 * Af

# Let's put all the carbon into the atmosphere
A <- Af + Bf
B <- 0

initial_state <- c(A = A, B = B)

# Time points for simulation
times <- seq(0, 150, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B")], type = "l", lty = 1, col = c("blue", "green"),
        xlab = "Time", ylab = "Carbon Amount",
        main = "Carbon Exchange between Atmosphere and Biosphere")

legend("topright", legend = c("Atmosphere", "Biosphere"), col = c("blue", "green"), lty = 1, bty = "n")

```

```{r 2boxModel-solution}

# Reference: 
# Avenhaus, R., S. Fenyi, and H. Frick. 1979. “Box Models for the CO2 Cycle of the Earth.” Environment International 2 (4): 379–85.

library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Exchange coefficients in yr-1 (Avenhaus et al., 1979; Table 1)
    k_ab <- 1/33 
    k_ba <- 1/40
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B - k_ab * A
    dB_dt <- k_ab * A - k_ba * B

    # Return the derivatives
    return(list(c(dA_dt, dB_dt)))
  })
}

# Initial state (Avenhaus et al., 1979; Table 1)
ya <- 1
yb <- 1.2 * ya

# Let's put all the carbon into the atmosphere
A <- ya + yb
B <- 0

initial_state <- c(A = A, B = B)

# Time points for simulation
times <- seq(0, 150, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B")], type = "l", lty = 1, col = c("blue", "green"),
        xlab = "Time (years)", ylab = "Carbon amount / atmospheric pool (-)")

legend("topright", legend = c("Atmosphere", "Biosphere"), col = c("blue", "green"), lty = 1, bty = "n")

```

### Exercise (2)

* Extend the model you built above by adding two additional pools, one for the ocean mixed layer and one for the ocean deep layer.

```{r 4boxModel, exercise = TRUE, exercise.eval = FALSE}

```

```{r 4boxModel-solution}

# Reference: 
# Avenhaus, R., S. Fenyi, and H. Frick. 1979. “Box Models for the CO2 Cycle of the Earth.” Environment International 2 (4): 379–85.

library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Exchange coefficients in yr-1 (Avenhaus et al., 1979; Table 1)
    k_ab <- 1/33  
    k_ba <- 1/40
    k_am <- 1/5 
    k_ma <- 1/6
    k_md <- 1/6.2
    k_dm <- 1/300
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B + k_ma * M - k_ab * A - k_am * A
    dB_dt <- k_ab * A - k_ba * B
    dM_dt <- k_am * A + k_dm * D - k_ma * M - k_md * M
    dD_dt <- k_md * M - k_dm * D

    # Return the derivatives
    return(list(c(dA_dt, dB_dt, dM_dt, dD_dt)))
  })
}

# Initial state

# The pool size is estimated by Craig

ya <- 1 
yb <- 1.21 * ya
ym <- 1.2 * ya
yd <- 58.06 * ya

A <- ya + yb + ym + yd
B <- 0
M <- 0
D <- 0

initial_state <- c(A = A, B = B, M = M, D = D)

# Time points for simulation
times <- seq(0, 300, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B", "M", "D")], type = "l", lty = 1, col = c("blue", "green", "magenta", "black"),
        xlab = "Time (years)", ylab = "Carbon amount / atm. pool (-)")

legend("right", legend = c("Atmosphere", "Biosphere", "Ocean Mixed Layer", "Ocean Deep Layer"), col = c("blue", "green", "magenta", "black"), lty = 1, bty = "n")

```

### Exercise (3)

* Next, run the same model, but with the pool sizes provided in our textbook in Figure 5.2 on page 75.

```{r 4boxModel02, exercise = TRUE, exercise.eval = FALSE}

```


```{r 4boxModel02-solution}

library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Exchange coefficients in yr-1 (Avenhaus et al., 1979; Table 1)
    k_ab <- 110/864
    k_ba <- 110/2150
    k_am <- 60/864
    k_ma <- 60/900
    k_md <- 100/900
    k_dm <- 100/37100
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B + k_ma * M - k_ab * A - k_am * A
    dB_dt <- k_ab * A - k_ba * B
    dM_dt <- k_am * A + k_dm * D - k_ma * M - k_md * M
    dD_dt <- k_md * M - k_dm * D

    # Return the derivatives
    return(list(c(dA_dt, dB_dt, dM_dt, dD_dt)))
  })
}

# Initial state

# The pool sizes

ya <- 864 
yb <- 2150
ym <- 900
yd <- 37100

A <- ya + yb + ym + yd
B <- 0
M <- 0
D <- 0

initial_state <- c(A = A, B = B, M = M, D = D)

# Time points for simulation
times <- seq(0, 500, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B", "M", "D")], type = "l", lty = 1, col = c("blue", "green", "magenta", "black"),
        xlab = "Time (years)", ylab = "Carbon amount (GtC)")

legend("right", legend = c("Atmosphere", "Biosphere", "Ocean Mixed Layer", "Ocean Deep Layer"), col = c("blue", "green", "magenta", "black"), lty = 1, bty = "n")

```


### Exercise (4)

* Next, let's emit 10,000 GtC to the atmosphere. Show how the carbon pool sizes changes in response. To do this, choose the currently estimated carbon pool sizes as initial values and add 10,000 GtC to the atmosphere.To contrast the resulting carbon pools, add horizontal lines of the current pool sizes to your graph:

```
abline(h = ya, col = "blue", lty = 2)
abline(h = yb, col = "green", lty = 2)
abline(h = ym, col = "magenta", lty = 2)
abline(h = yd, col = "black", lty = 2)
```

* Enhance the readability of your Figure by reducing the y-axis range to a maximum value of 7000 GtC:

```
ylim = c(0, 7000)
```

* Change the length of the time axis to evaluate whether the amount of carbon in the atmosphere will ever return to its initial value.

* What is the main limitation of this approach?

```{r 4boxModel03, exercise = TRUE, exercise.eval = FALSE}

```


```{r 4boxModel03-solution}

library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Exchange coefficients in yr-1 (Avenhaus et al., 1979; Table 1)
    k_ab <- 110/864
    k_ba <- 110/2150
    k_am <- 60/864
    k_ma <- 60/900
    k_md <- 100/900
    k_dm <- 100/37100
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B + k_ma * M - k_ab * A - k_am * A
    dB_dt <- k_ab * A - k_ba * B
    dM_dt <- k_am * A + k_dm * D - k_ma * M - k_md * M
    dD_dt <- k_md * M - k_dm * D

    # Return the derivatives
    return(list(c(dA_dt, dB_dt, dM_dt, dD_dt)))
  })
}

# Initial state

# The pool sizes

ya <- 864 
yb <- 2150
ym <- 900
yd <- 37100

# emissions (GtC)
ye <- 10000

A <- ya + ye
B <- yb
M <- ym
D <- yd

initial_state <- c(A = A, B = B, M = M, D = D)

# Time points for simulation
times <- seq(0, 500, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B", "M", "D")], type = "l", lty = 1, col = c("blue", "green", "magenta", "black"),
        xlab = "Time (years)", ylab = "Carbon amount (GtC)", ylim = c(0,3000))

abline(h = ya, col = "blue", lty = 2)
abline(h = yb, col = "green", lty = 2)
abline(h = ym, col = "magenta", lty = 2)
abline(h = yd, col = "black", lty = 2)


legend("topright", legend = c("Atmosphere", "Biosphere", "Ocean Mixed Layer", "Ocean Deep Layer"), col = c("blue", "green", "magenta", "black"), lty = 1, bty = "n")

```
