## The Carbon Cycle
### Exercise (1)
* Make a 2-box model that simulates the exchange of carbon between the atmosphere the biosphere. For the inital time step, put all the carbon into the atmosphere and run the model until it reaches equilibrium. Use the following reference for building your model: Avenhaus, R., S. Fenyi, and H. Frick. 1979. “Box Models for the CO2 Cycle of the Earth.” Environment International 2 (4): 379–85.

```{r 2boxModel, exercise = TRUE, exercise.eval = FALSE}

# Reference: 
# Sawyer, J. S. 1972. “Man-Made Carbon Dioxide and the ‘Greenhouse’ Effect.” Nature 239 (5366): 23–26.

# Install and load the deSolve package
# install.packages("deSolve")
library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Parameters
    k_ab <- 1/33  # Transfer rate from atmosphere to biosphere measured in yr-1
    k_ba <- 1/40 # Transfer rate from biosphere to atmosphere measured in yr-1
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- ___
    dB_dt <- ___

    # Return the derivatives
    return(list(c(dA_dt, dB_dt)))
  })
}

# Initial state
Af <- 1
Bf <- 1.2 * Af

# Let's put all the carbon into the atmosphere
A <- Af + Bf
B <- 0

initial_state <- c(A = A, B = B)

# Time points for simulation
times <- seq(0, 150, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B")], type = "l", lty = 1, col = c("blue", "green"),
        xlab = "Time", ylab = "Carbon Amount",
        main = "Carbon Exchange between Atmosphere and Biosphere")

legend("topright", legend = c("Atmosphere", "Biosphere"), col = c("blue", "green"), lty = 1, bty = "n")

```

```{r 2boxModel-solution}

# Reference: 
# Sawyer, J. S. 1972. “Man-Made Carbon Dioxide and the ‘Greenhouse’ Effect.” Nature 239 (5366): 23–26.

# Install and load the deSolve package
# install.packages("deSolve")
library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Parameters
    k_ab <- 1/33  # Transfer rate from atmosphere to biosphere measured in yr-1
    k_ba <- 1/40 # Transfer rate from biosphere to atmosphere measured in yr-1
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B - k_ab * A
    dB_dt <- k_ab * A - k_ba * B

    # Return the derivatives
    return(list(c(dA_dt, dB_dt)))
  })
}

# Initial state
Af <- 1
Bf <- 1.2 * Af

# Let's put all the carbon into the atmosphere
A <- Af + Bf
B <- 0

initial_state <- c(A = A, B = B)

# Time points for simulation
times <- seq(0, 150, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B")], type = "l", lty = 1, col = c("blue", "green"),
        xlab = "Time", ylab = "Carbon Amount",
        main = "Carbon Exchange between Atmosphere and Biosphere")

legend("topright", legend = c("Atmosphere", "Biosphere"), col = c("blue", "green"), lty = 1, bty = "n")

```

### Exercise (2)

* Extend the model you built above by adding two additional pools, one for the ocean mixed layer and one for the ocean deep layer.

```{r 4boxModel, exercise = TRUE, exercise.eval = FALSE}

```

```{r 4boxModel-solution}

# Reference: 
# Sawyer, J. S. 1972. “Man-Made Carbon Dioxide and the ‘Greenhouse’ Effect.” Nature 239 (5366): 23–26.

# Install and load the deSolve package
# install.packages("deSolve")
library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Parameters
    k_ab <- 1/33  # Transfer rate from atmosphere to biosphere measured in yr-1
    k_ba <- 1/40 # Transfer rate from biosphere to atmosphere measured in yr-1
    k_am <- 1/5 
    k_ma <- 1/6
    k_md <- 1/4
    k_dm <- 1/300
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B + k_ma * M - k_ab * A - k_am * A
    dB_dt <- k_ab * A - k_ba * B
    dM_dt <- k_am * A + k_dm * D - k_ma * M - k_md * M
    dD_dt <- k_md * M - k_dm * D

    # Return the derivatives
    return(list(c(dA_dt, dB_dt, dM_dt, dD_dt)))
  })
}

# Initial state

# The pool size is estimated by Craig

Af <- 1 
Bf <- 1.2 * Af
Mf <- 1.2 * Af
Df <- 58 * Af

A <- Af + Bf + Mf + Df
B <- 0
M <- 0
D <- 0

initial_state <- c(A = A, B = B, M = M, D = D)

# Time points for simulation
times <- seq(0, 300, by = 0.1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B", "M", "D")], type = "l", lty = 1, col = c("blue", "green", "magenta", "black"),
        xlab = "Time", ylab = "Carbon Amount",
        main = "Carbon Exchange between Atmosphere, Biosphere, and Ocean")

legend("right", legend = c("Atmosphere", "Biosphere", "Ocean Mixed Layer", "Ocean Deep Layer"), col = c("blue", "green", "magenta", "black"), lty = 1, bty = "n")

```

### Exercise (3)
* Next, run the same model, but adding fossil fuel emissions.

```{r 4boxModelwithFossilFuelEmissions, exercise = TRUE, exercise.eval = FALSE}

```


```{r 4boxModelwithFossilFuelEmissions-solution}

# Reference: 
# Sawyer, J. S. 1972. “Man-Made Carbon Dioxide and the ‘Greenhouse’ Effect.” Nature 239 (5366): 23–26.

# Install and load the deSolve package
# install.packages("deSolve")
library(deSolve)

# Define the box model function
carbon_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    # Parameters
    k_ab <- 1/33  # Transfer rate from atmosphere to biosphere measured in yr-1
    k_ba <- 1/40 # Transfer rate from biosphere to atmosphere measured in yr-1
    k_am <- 1/5 
    k_ma <- 1/6
    k_md <- 1/4
    k_dm <- 1/300
    
    # ODEs: Flux = sources minus sinks
    dA_dt <- k_ba * B + k_ma * M - k_ab * A - k_am * A + N
    dB_dt <- k_ab * A - k_ba * B
    dM_dt <- k_am * A + k_dm * D - k_ma * M - k_md * M
    dD_dt <- k_md * M - k_dm * D

    # Return the derivatives
    return(list(c(dA_dt, dB_dt, dM_dt, dD_dt)))
  })
}

# Initial state

# Convert current annual emissions from Gt to mol
emissions <- 36.8 # current CO2 emissions:Gt
emissions <- emissions * 10^9 * 10^6 # g CO2
molarMassCO2 <- 44.01 # g/mol
emissions <- emissions / molarMassCO2

N <- emissions

N <- 0

A <- 51 * 10^15
B <- 62.2 * 10^15
M <- 61.7 * 10^15
D <- 2985.4 * 10^15


A + B + M + D

initial_state <- c(A = A, B = B, M = M, D = D)

# Time points for simulation
times <- seq(0, 1000, by = 1)

# Parameters (not used in this simple example)
parameters <- list()

# Solve the ODEs
output <- ode(y = initial_state, times = times, func = carbon_model, parms = parameters)

output <- data.frame(output)

# Plot the results
matplot(output$time, output[, c("A", "B", "M", "D")], type = "l", lty = 1, col = c("blue", "green", "magenta", "black"),
        xlab = "Time", ylab = "Carbon Amount",
        main = "Carbon Exchange between Atmosphere, Biosphere, and Ocean")

legend("right", legend = c("Atmosphere", "Biosphere", "Ocean Mixed Layer", "Ocean Deep Layer"), col = c("blue", "green", "magenta", "black"), lty = 1, bty = "n")

```