## Past and Future Changes

### Exercise (1)
The code below plots a time series of the daily mean incoming solar radiation (W m^-2^) at the top of the atmosphere (TOA) during summer solstice (Jun 20) at the latitude of 65 Degrees North. The horizontal line gives the corresponding value for the year 1950 (i.e. year zero), which I obtained from [here](https://data.giss.nasa.gov/modelE/ar5plots/srlocat.html). The vertical line gives the time of the present (2024). 

The government has asked you to evaluate to what extent the Milankovitch cycle affects our current climate. To prepare your response, answer the following questions. 

* By how much has the insolation changed during the Holocene?
* Did the insolation increase or decrease since the onset of the industrial revolution?
* Looking into the future, how will the Milankovitch cycle affect warming over the next 1000 years?
* How large is this affect compared to a radiative forcing of 8.5 W m^-2^?

You can answer these questions by running the code below for different time periods defined as ``` year <- seq(from = -400000, to = 0, by = 1000) ```

```{r Milankovitch, exercise = TRUE, exercise.eval = FALSE}
library(palinsol)

year <- seq(from = -10^6, to = 0, by = 1000) 
# year <- seq(from = -10000, to = 0, by = 10) 

latitude <- 65 * pi/180 # (radians)

insolation <- function(times, astrosol=ber78,...)
sapply(times, function(tt) Insol(orbit=astrosol(tt), lon = pi/2, lat = latitude))

# Daily mean incoming solar radiation at TOA (W/m2) 
isl <- insolation(year, ber78)

plot(year, isl, typ='l', 
     main = "Daily mean solar radiation for June 20 (Solstice)",
     xlab = "Years",
     ylab = "Insolation at TOA (W/m2)")
legend("topleft", paste("Latitude =", latitude * 180/pi, "N"), bty = "n")
abline(h = 480.05, col = "grey")
abline(v=2024-1950)

```


### Exercise (2)

In this exercise you will run a simple global climate model called [Hector](https://jgcri.github.io/hector/index.html). Hector is the default carbon cycle module for the integrated assessment model GCAM (Global Change Assessment Model). The section below will show you how to run Hector. 

* Run the model

```{r hector01, exercise = TRUE, exercise.eval = FALSE}
library(hector)
ini_file <- system.file("input/hector_ssp585.ini", package = "hector")

# Initialize a Hector core. The core is a self-contained object that contains information about all of Hector’s inputs and outputs
core <- newcore(ini_file)

# Run the model
run(core)

# Obtain the results
results_default <- fetchvars(core, 1950:2300)

# Have a look at the results
head(results_default)
```

* Plot model output

```{r hector02, exercise = TRUE, exercise.eval = FALSE, fig.width = 6, fig.height = 5}

library(hector)
ini_file <- ___

# Initialize a Hector core. The core is a self-contained object that contains information about all of Hector’s inputs and outputs
core <- ___

# Run the model
run(___)

# Obtain the results
results_default <- ___

# Rename your results "df" (data frame). This is just for giving it a shorter name
df <- results_default

# There are four output variables
categories <- unique(df$variable)

# Set up a layout for multiple plots
par(mfrow = c(2,2), mar = c(4,4,1,1))

# Loop through each category and create a plot for each category
for (i in 1:length(categories)) {
  
  # Subset data frame 
  data <- df[df$variable == categories[i], ]
  unit <- unique(data$units)
  # Plot data
  plot(x = data$year, y = data$value, ylab = paste(categories[i], unit), xlab = "Year", type = "l")
  
```


```{r hector02-solution, fig.width = 6, fig.height = 5}

library(hector)
ini_file <- system.file("input/hector_ssp585.ini", package = "hector")

# Initialize a Hector core. The core is a self-contained object that contains information about all of Hector’s inputs and outputs
core <- newcore(ini_file)

# Run the model
run(core)

# Obtain the results
results_default <- fetchvars(core, 1950:2300)

# Rename your results "df" (data frame). This is just for giving it a shorter name
df <- results_default

# There are four output variables
categories <- unique(df$variable)

# Set up a layout for multiple plots
par(mfrow = c(2,2), mar = c(4,4,1,1))

# Loop through each category and create a plot for each category
for (i in 1:length(categories)) {
  
  # Subset data frame 
  data <- df[df$variable == categories[i], ]
  unit <- unique(data$units)
  # Plot data
  plot(x = data$year, y = data$value, ylab = paste(categories[i], unit), xlab = "Year", type = "l")
}

```

### Exercise (3)

The model accounts for the CO~2~ fertilization effect using a multiplier $f$:

$$ f = 1 + \beta_i \left( \log \left( \frac{C_{atm}}{C0}\right) \right) $$

* Let's increase the parameter $\beta_i$ from its default value to 0.8 and run the model again

```{r hector03, exercise = TRUE, exercise.eval = FALSE}
library(hector)
ini_file <- system.file("input/hector_ssp585.ini", package = "hector")

# Initialize a Hector core. The core is a self-contained object that contains information about all of Hector’s inputs and outputs
core <- newcore(ini_file)

# Run the model
run(core)

# The current value for beta is
beta <- fetchvars(core, NA, BETA())

# Let's increase this to 0.8
setvar(core, NA, BETA(), 0.80, "(unitless)")

# Reset the core with the new value
reset(core)

# run the model
run(core)

# Get the results
results_beta <- fetchvars(core, 1950:2300)
```

* Show the impact by plotting the results from the default run and from the new run. How does the increase in $\beta_i$ affect the atmospheric CO~2~ concentration, radiative forcing, and global mean surface temperature? 

```{r hector04, exercise = TRUE, exercise.eval = FALSE, fig.width = 6, fig.height = 5}

# add code here

df01 <- results_default
df02 <- results_beta

categories <- unique(df01$variable)

# Set up a layout for multiple plots
par(mfrow = c(2,2), mar = c(4,4,1,1))

# Loop through each category and create a plot for each category
for (i in 1:length(categories)) {
  
  # Subset data frame 
  data <- df01[df01$variable == categories[i], ]
  unit <- unique(data$units)
  # Plot data
  plot(x = data$year, y = data$value, ylab = paste(categories[i], unit), xlab = "Year", type = "l")
  
  data <- df02[df02$variable == categories[i], ]
  lines(x = data$year, y = data$value, type = "l", col = "red")
  abline(v = 2100)
    
}

```

```{r hector04-solution, fig.width = 6, fig.height = 5}
library(hector)
ini_file <- system.file("input/hector_ssp585.ini", package = "hector")

# Initialize a Hector core.
core <- newcore(ini_file)

# Run the model
run(core)

# Obtain the results
results_default <- fetchvars(core, 1950:2300)

# The current value for beta is
beta <- fetchvars(core, NA, BETA())

# Let's increase this to 0.5
setvar(core, NA, BETA(), 0.80, "(unitless)")

# Reset the core with the new value
reset(core)

# run the model
run(core)

# Get the results
results_beta <- fetchvars(core, 1950:2300)

df01 <- results_default
df02 <- results_beta

categories <- unique(df01$variable)

# Set up a layout for multiple plots
par(mfrow = c(2,2), mar = c(4,4,1,1))

# Loop through each category and create a plot for each category
for (i in 1:length(categories)) {
  
  # Subset data frame 
  data <- df01[df01$variable == categories[i], ]
  unit <- unique(data$units)
  # Plot data
  plot(x = data$year, y = data$value, ylab = paste(categories[i], unit), xlab = "Year", type = "l")
  
  data <- df02[df02$variable == categories[i], ]
  lines(x = data$year, y = data$value, type = "l", col = "red")
  abline(v = 2100)
    
}

```

### Exercise (4)

The government has asked you to conduct a study that shows how global mean surface temperature would respond if we immediately stopped all CO~2~ emissions. Hector simulates atmospheric CO~2~ concentrations, and radiative forcings, and global mean surface temperature based on prescribed emissions. To respond to the government's request, you need to change the prescribed emissions and simulate the impacts. To do this 

* Type ```ini_file``` into your R console to see the location of your initialization file

* You should obtain a directory similar to this one:

```
C:/Users/seile/AppData/Local/R/win-library/4.2/hector/input/hector_ssp585.ini
```

* Copy the file and give it a new name:

```
C:/Users/seile/AppData/Local/R/win-library/4.2/hector/input/hector_ENSC841.ini
```

* Open the file and replace all

```
ssp585_emiss-constraints_rf.csv
```

with 

```
ENSC841.csv
```

* Go the the tables folder and copy ```ssp585_emiss-constraints_rf.csv``` to ```ENSC841.csv```

* Open ```ENSC841.csv``` in Excel and have a look at the ```ffi_emissions``` column. This column gives the emissions in PgC per year. Set these values to zero starting from 2024 onward.

* Run the model withe the new initialization file

```ini_file <- system.file("input/hector_ENSC841.ini", package = "hector")```

* Would this be sufficient to keep temperatures below 1.5 Degrees Celsius by 2100?

```{r hector05, exercise = TRUE, exercise.eval = FALSE, fig.width = 6, fig.height = 5}

```

```{r hector05-solution, fig.width = 6, fig.height = 5}
library(hector)
ini_file <- system.file("input/hector_ENSC841.ini", package = "hector")

# Initialize a Hector core.
core <- newcore(ini_file)

# Run the model
run(core)

# Obtain the results
df <- fetchvars(core, 1950:2300)

# There are four output variables
categories <- unique(df$variable)

# Set up a layout for multiple plots
par(mfrow = c(2,2), mar = c(4,4,1,1))

# Loop through each category and create a plot for each category
for (i in 1:length(categories)) {
  
  # Subset data frame 
  data <- df[df$variable == categories[i], ]
  unit <- unique(data$units)
  # Plot data
  plot(x = data$year, y = data$value, ylab = paste(categories[i], unit), xlab = "Year", type = "l")
  abline(v=2024, lty = 2)
  abline(v=2100, lty = 2)
}
```

* Repeat the experiment above, but also reduce methane (CH~4~) emissions from 2024 onward to zero

* Next, also set nitrous oxide (N~2~O) emissions from 2024 onward to zero

* How do these changes affect global temperature? 
