## Mitigation Policies


### Exercise (1)

Using the data from the table below, make a Figure that shows how the sum of the costs associated with (a) reducing emissions and (b) the carbon tax of 4 CAD per tonne of carbon change as emissions are reduced from 0 to 7 tonnes of carbon.


| Emissions reduced (tC) | Emissions (tC)| Marginal Costs (CAD)| Total Costs CAD|
|:---:|:---:|:---:|:---:|
| 0  | 10  | 0 | 0 |
| 1  | 9   | 1 | 1 |
| 2  | 8   | 2 | 3 |
| 3  | 7   | 3 | 6 |
| 4  | 6   | 4 | 10 |
| 5  | 5   | 5 | 15 |
| 6  | 4   | 6 | 21 |
| 7  | 3   | 7 | 28 |


```{r carbonTax, exercise = TRUE, exercise.eval = FALSE}
```

```{r carbonTax-solution}

carbonTax <- 4 # ($ per tC) 
reduction <- seq(0,7, 1)
emission <- 10 - reduction
marginal <- seq(0,7, 1)
total <- cumsum(marginal)
tax <- emission * carbonTax
TotalPlusTax <- total + tax 

par(mfrow = c(1,1), mar = c(5,5,1,1))
plot(reduction, TotalPlusTax, type = "l",
     xlab = "Emissions reduced (tC)",
     ylab = "Costs of reductions + Tax (CAD)")

```
### Exercise (2)

In this exercise you will explore climate change mitigation pathways that are consistent with the 2015 Paris Agreement. The pathways are used in the [Special Report on Global Warming of 1.5$^{\circ}$C (SR15)](https://www.ipcc.ch/site/assets/uploads/sites/2/2022/06/SR15_Full_Report_LR.pdf) by the Intergovernmental Panel on Climate Change (IPCC) published in 2018. The datasets are documented [here](https://data.ene.iiasa.ac.at/iamc-1.5c-explorer/#/docs). To download the data visit the [Integrated Assessment Modeling Consortium scenario explorer](https://data.ene.iiasa.ac.at/iamc-1.5c-explorer/#/downloads). Download the following spreadsheet to your computer:

```
iamc15_scenario_data_world_r2.0.xlsx
```

Copy-paste the code below into the R console. Make sure you edit the path where R can find your spreadsheet.

```
rm(list = ls())
library(mipplot)
library(tidyverse)
data <- read_iamc_xlsx("C:/Users/seile/Documents/queensu/teaching/classes/ENSC841/data/IAMC/iamc15_scenario_data_world_r2.0.xlsx", sheet = 2)

mipplot_interactive_line(data)

```

If all went well, a new window appears. Make the following selections in the dropdown lists:

* ```region: World ```
* ```variable: Emissions|CO2 ```
* ```model: REMIND-MAgPIE 1.7-3.0 ```
* ```scenario: SMP_1p5C_Def, SMP_2C_Def, and SMP_REF_Def ```

The three scenarios are described [here](https://iopscience.iop.org/article/10.1088/1748-9326/aac3ec/pdf) and correspond to the following carbon budget and mitigation measures:

* ```SMP_1p5C_Def```: Budget of 400Gt CO2 2011-2100 with default policy of comprehensive carbon pricing only
* ```SMP_2C_Def```: Budget of 1000Gt CO2 2011-2100 with default policy of comprehensive carbon pricing only
* ```SMP_REF_Def```: No-policy baseline

Hit the ```Apply Changes``` button to visualize the data you selected. Save the images and create a small slide deck that answers the following questions: 

* How do the three pathways differ? 
* When do emissions need to become negative for reaching the 1.5 Degrees target?
* Show the probability that global mean surface temperature will exceed 1.5 degrees (Select ```variable: AR5 climate diagnostics|Temperature|Exceedance Probability|1.5C|MAGICC6 ``` and hit the ```Apply Changes``` button)
* What are the projections for the carbon price and how do the values compare to Canada's carbon tax? (Select ```Price|Carbon```)
* How does the carbon price change over time and 
* How do we generate negative emissions? To answer this question, explore the role of forests (```Land Cover|Forest```, ```Land Cover|Forest|Natural Forest```)
* How does the energy supply changes over time (```Capacity|Electric|Solar```)?
* How important is carbon capture and storage (```Carbon Sequestration|CCS```)?
* Explore other variables that you find interesting. Note that most scenarios are model-specific, and not all variables are available for all models



```{r eval = FALSE}

carbonTax <- seq(0,6,2) # ($ per tC) 
reduction <- seq(0,7,1)

my.fun <- function(reduction, carbonTax){

emission <- 10 - reduction
marginal <- seq(0,7, 1)
total <- cumsum(marginal)
tax <- emission * carbonTax
TotalPlusTax <- total + tax
return(TotalPlusTax)
}


data <- outer(reduction, carbonTax, my.fun)

my.xlab <- "Emissions reduced (tC)"
my.ylab <- "Costs of reductions + Tax (CAD)"

my.col <- rev(viridis::plasma(n = ncol(data)))

par(mar = c(5,5,1,1))
matplot(x = reduction, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("Carbon Tax =", carbonTax, "CAD/tC", sep = " ")
legend("bottomright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r eval = FALSE}
# Plant A
reduction <- seq(0,7, 1)
carbonTax <- 4 # ($ per tC) 
emission <- 10 - reduction
marginal <- seq(0,7, 1)
total <- cumsum(marginal)
tax <- emission * carbonTax
plantA <- total + tax 

# Plant B
reduction <- seq(0,7, 1)
carbonTax <- 4 # ($ per tC) 
emission <- 10 - reduction
marginal <- seq(0,7, 1) * 2
total <- cumsum(marginal)
tax <- emission * carbonTax
plantB <- total + tax 



par(mfrow = c(1,1), mar = c(5,5,1,1))
plot(reduction, plantA, type = "l", ylim = c(30,70),
     xlab = "Emissions reduced (tC)",
     ylab = "Costs of reductions + Tax (CAD)")

lines(reduction, plantB, col = "red")

my.col <- c("black", "red")

legend.text <- paste("Carbon Tax =", carbonTax, "CAD/tC", sep = " ")
legend("topleft", legend = c("Plant A", "Pant B (marginal costs 2 x that of Plant A)"), pch = 16, col = my.col, bty = "n")

```

