## Mitigation Policies

### Exercise (1)

Using the data from the table below, make a Figure that shows how the sum of the costs associated with (a) reducing emissions and (b) the carbon tax of 4 CAD per tonne of carbon change as emissions are reduced from 0 to 7 tonnes of carbon.


| Emissions reduced (tC) | Emissions (tC)| Marginal Costs (CAD)| Total Costs CAD|
|:---:|:---:|:---:|:---:|
| 0  | 10  | 0 | 0 |
| 1  | 9   | 1 | 1 |
| 2  | 8   | 2 | 3 |
| 3  | 7   | 3 | 6 |
| 4  | 6   | 4 | 10 |
| 5  | 5   | 5 | 15 |
| 6  | 4   | 6 | 21 |
| 7  | 3   | 7 | 28 |


```{r carbonTax, exercise = TRUE, exercise.eval = FALSE}
```

```{r carbonTax-solution}

carbonTax <- 4 # ($ per tC) 
reduction <- seq(0,7, 1)
emission <- 10 - reduction
marginal <- seq(0,7, 1)
total <- cumsum(marginal)
tax <- emission * carbonTax
TotalPlusTax <- total + tax 

par(mfrow = c(1,1), mar = c(5,5,1,1))
plot(reduction, TotalPlusTax, type = "l",
     xlab = "Emissions reduced (tC)",
     ylab = "Costs of reductions + Tax (CAD)")

```
### Exercise (2)

In this exercise you will explore climate change mitigation pathways that are consistent with the 2015 Paris Agreement. The pathways are used in the [Special Report on Global Warming of 1.5$^{\circ}$C (SR15)](https://www.ipcc.ch/site/assets/uploads/sites/2/2022/06/SR15_Full_Report_LR.pdf) by the Intergovernmental Panel on Climate Change (IPCC) published in 2018. The datasets are documented [here](https://data.ene.iiasa.ac.at/iamc-1.5c-explorer/#/docs). To download the data visit the [Integrated Assessment Modeling Consortium scenario explorer](https://data.ene.iiasa.ac.at/iamc-1.5c-explorer/#/downloads). Download the following spreadsheet to your computer:

```
iamc15_scenario_data_world_r2.0.xlsx
```

Run the code chunk below. Make sure you edit the path where R can find your ```iamc15_scenario_data_world_r2.0.xlsx``` spreadsheet.


```{r, IAMC15a, exercise = TRUE}

options(warn=-1)

rm(list = ls())
library(mipplot)
library(tidyverse)

myFile <- file.path("data", "iamc15_scenario_data_world_r2.0.xlsx") 
data <- read_iamc_xlsx(myFile, sheet = 2)
data_subset <- data %>%
filter( model %in% c("REMIND-MAgPIE 1.7-3.0") ) %>%
filter(2000 <= period) %>%
filter(period <= 2100)

mipplot_line(
data_subset,
variable = c("Emissions|CO2"),
scenario = c("SMP_1p5C_Def", "SMP_2C_Def", "SMP_REF_Def"),
region = c("World"),
legend = TRUE,
axis_year_text_angle = 0,
language = 'en')
```

The three scenarios selected above are described [here](https://iopscience.iop.org/article/10.1088/1748-9326/aac3ec/pdf) and correspond to the following carbon budget and mitigation measures:

* ```SMP_1p5C_Def```: Budget of 400Gt CO2 2011-2100 with default policy of comprehensive carbon pricing only
* ```SMP_2C_Def```: Budget of 1000Gt CO2 2011-2100 with default policy of comprehensive carbon pricing only
* ```SMP_REF_Def```: No-policy baseline

You can choose different models, variables, and scenarios, using the respective names given in the boxes below. Note that not all combinations are possible, e.g. most scenarios are model-specific.


Answers the following questions for the model ```REMIND-MAgPIE 1.7-3.0``` and the scenarios ```SMP_1p5C_Def```, ```SMP_2C_Def```, and ```SMP_REF_Def```. Save the images and present a short slide deck that illustrates your answers.

* How do the three pathways differ? 
* When do emissions need to become negative for reaching the 1.5 Degrees target?
* Show the probability that global mean surface temperature will exceed 1.5 degrees. To answer this question, you need to replace the variable CO2 emissions (```Emissions|CO2```) with the probability that our selected scenarios will exceed our 1.5 Degrees Celsius warming target (```AR5 climate diagnostics|Temperature|Exceedance Probability|1.5 °C|MAGICC6```):

```{r IAMC15b, exercise = TRUE}

options(warn=-1)

rm(list = ls())
library(mipplot)
library(tidyverse)

myFile <- file.path("data", "iamc15_scenario_data_world_r2.0.xlsx") 
data <- read_iamc_xlsx(myFile, sheet = 2)

data_subset <- data %>%
filter( model %in% c("REMIND-MAgPIE 1.7-3.0") ) %>%
filter(2000 <= period) %>%
filter(period <= 2100)

mipplot_line(
data_subset,
variable = c("AR5 climate diagnostics|Temperature|Exceedance Probability|1.5 °C|MAGICC6"),
scenario = c("SMP_1p5C_Def", "SMP_2C_Def", "SMP_REF_Def"),
region = c("World"),
legend = TRUE,
axis_year_text_angle = 0,
language = 'en')
```

* What are the projections for the carbon price and how do the values compare to Canada's carbon tax? (variable: ```Price|Carbon```)
* How does the carbon price change over time?
* How do we generate negative emissions? To answer this question, explore the role of forests (variable: ```Land Cover|Forest```, ```Land Cover|Forest|Natural Forest```)
* How does the energy supply changes over time (variable: ```Capacity|Electricity|Solar```)?
* How important is carbon capture and storage (variable: ```Carbon Sequestration|CCS```)?
* Explore other variables that you find interesting. Recall that most scenarios are model-specific, and not all variables are available for all models

#### Models

```
 1 AIM/CGE 2.0                                 
 2 AIM/CGE 2.1                                 
 3 C-ROADS-5.005                               
 4 GCAM 4.2                                    
 5 GENeSYS-MOD 1.0                             
 6 IEA Energy Technology Perspective Model 2017
 7 IEA World Energy Model 2017                 
 8 IMAGE 3.0.1                                 
 9 IMAGE 3.0.2                                 
10 MERGE-ETL 6.0                               
11 MESSAGE V.3                                 
12 MESSAGE-GLOBIOM 1.0                         
13 MESSAGEix-GLOBIOM 1.0                       
14 POLES ADVANCE                               
15 POLES CD-LINKS                              
16 POLES EMF33                                 
17 REMIND 1.5                                  
18 REMIND 1.7                                  
19 REMIND-MAgPIE 1.5                           
20 REMIND-MAgPIE 1.7-3.0                       
21 Reference                                   
22 Shell World Energy Model 2018               
23 WITCH-GLOBIOM 3.1                           
24 WITCH-GLOBIOM 4.2                           
25 WITCH-GLOBIOM 4.4 
```

#### Variables

```
  1 AR5 climate diagnostics|Concentration|CO2|FAIR|MED                                 ppm                
  2 AR5 climate diagnostics|Concentration|CO2|MAGICC6|MED                              ppm                
  3 AR5 climate diagnostics|Forcing|Aerosol|Direct|MAGICC6|MED                         W/m2               
  4 AR5 climate diagnostics|Forcing|Aerosol|MAGICC6|MED                                W/m2               
  5 AR5 climate diagnostics|Forcing|Aerosol|Total|FAIR|MED                             W/m2               
  6 AR5 climate diagnostics|Forcing|Aerosol|Total|MAGICC6|MED                          W/m2               
  7 AR5 climate diagnostics|Forcing|CH4|FAIR|MED                                       W/m2               
  8 AR5 climate diagnostics|Forcing|CH4|MAGICC6|MED                                    W/m2               
  9 AR5 climate diagnostics|Forcing|CO2|FAIR|MED                                       W/m2               
 10 AR5 climate diagnostics|Forcing|CO2|MAGICC6|MED                                    W/m2               
 11 AR5 climate diagnostics|Forcing|F-Gases|MAGICC6|MED                                W/m2               
 12 AR5 climate diagnostics|Forcing|FAIR|MED                                           W/m2               
 13 AR5 climate diagnostics|Forcing|Kyoto Gases|MAGICC6|MED                            W/m2               
 14 AR5 climate diagnostics|Forcing|MAGICC6|MED                                        W/m2               
 15 AR5 climate diagnostics|Forcing|Montreal Protocol (orig.) gases|MAGICC6|MED        W/m2               
 16 AR5 climate diagnostics|Forcing|N2O|FAIR|MED                                       W/m2               
 17 AR5 climate diagnostics|Forcing|N2O|MAGICC6|MED                                    W/m2               
 18 AR5 climate diagnostics|Forcing|OtherGHGs|FAIR|MED                                 W/m2               
 19 AR5 climate diagnostics|Forcing|Tropospheric Ozone|FAIR|MED                        W/m2               
 20 AR5 climate diagnostics|Forcing|Tropospheric Ozone|MAGICC6|MED                     W/m2               
 21 AR5 climate diagnostics|Temperature|Exceedance Probability|1.0 °C|FAIR             -                  
 22 AR5 climate diagnostics|Temperature|Exceedance Probability|1.0 °C|MAGICC6          -                  
 23 AR5 climate diagnostics|Temperature|Exceedance Probability|1.5 °C|FAIR             -                  
 24 AR5 climate diagnostics|Temperature|Exceedance Probability|1.5 °C|MAGICC6          -                  
 25 AR5 climate diagnostics|Temperature|Exceedance Probability|2.0 °C|FAIR             -                  
 26 AR5 climate diagnostics|Temperature|Exceedance Probability|2.0 °C|MAGICC6          -                  
 27 AR5 climate diagnostics|Temperature|Exceedance Probability|2.5 °C|FAIR             -                  
 28 AR5 climate diagnostics|Temperature|Exceedance Probability|2.5 °C|MAGICC6          -                  
 29 AR5 climate diagnostics|Temperature|Exceedance Probability|3.0 °C|FAIR             -                  
 30 AR5 climate diagnostics|Temperature|Exceedance Probability|3.0 °C|MAGICC6          -                  
 31 AR5 climate diagnostics|Temperature|Exceedance Probability|3.5 °C|FAIR             -                  
 32 AR5 climate diagnostics|Temperature|Exceedance Probability|3.5 °C|MAGICC6          -                  
 33 AR5 climate diagnostics|Temperature|Exceedance Probability|4.0 °C|FAIR             -                  
 34 AR5 climate diagnostics|Temperature|Exceedance Probability|4.0 °C|MAGICC6          -                  
 35 AR5 climate diagnostics|Temperature|Global Mean|FAIR|Expected value                °C                 
 36 AR5 climate diagnostics|Temperature|Global Mean|FAIR|Expected value|Non CO2        °C                 
 37 AR5 climate diagnostics|Temperature|Global Mean|FAIR|MED                           °C                 
 38 AR5 climate diagnostics|Temperature|Global Mean|FAIR|MED|Non CO2                   °C                 
 39 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P5                            °C                 
 40 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P5|Non CO2                    °C                 
 41 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P66                           °C                 
 42 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P66|Non CO2                   °C                 
 43 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P90                           °C                 
 44 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P90|Non CO2                   °C                 
 45 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P95                           °C                 
 46 AR5 climate diagnostics|Temperature|Global Mean|FAIR|P95|Non CO2                   °C                 
 47 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|Expected value             °C                 
 48 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|Expected value|Non CO2     °C                 
 49 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|MED                        °C                 
 50 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|P5                         °C                 
 51 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|P66                        °C                 
 52 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|P90                        °C                 
 53 AR5 climate diagnostics|Temperature|Global Mean|MAGICC6|P95                        °C                 
 54 Capacity|Electricity|Biomass                                                       GW                 
 55 Capacity|Electricity|Coal                                                          GW                 
 56 Capacity|Electricity|Gas                                                           GW                 
 57 Capacity|Electricity|Geothermal                                                    GW                 
 58 Capacity|Electricity|Hydro                                                         GW                 
 59 Capacity|Electricity|Nuclear                                                       GW                 
 60 Capacity|Electricity|Solar                                                         GW                 
 61 Capacity|Electricity|Solar|PV                                                      GW                 
 62 Capacity|Electricity|Wind                                                          GW                 
 63 Capacity|Electricity|Wind|Onshore                                                  GW                 
 64 Carbon Sequestration|CCS                                                           Mt CO2/yr          
 65 Carbon Sequestration|CCS|Biomass                                                   Mt CO2/yr          
 66 Carbon Sequestration|CCS|Biomass|Energy|Supply                                     Mt CO2/yr          
 67 Carbon Sequestration|CCS|Biomass|Energy|Supply|Electricity                         Mt CO2/yr          
 68 Carbon Sequestration|CCS|Biomass|Energy|Supply|Liquids                             Mt CO2/yr          
 69 Carbon Sequestration|CCS|Fossil                                                    Mt CO2/yr          
 70 Carbon Sequestration|CCS|Fossil|Energy|Demand|Industry                             Mt CO2/yr          
 71 Carbon Sequestration|CCS|Fossil|Energy|Supply                                      Mt CO2/yr          
 72 Carbon Sequestration|CCS|Fossil|Energy|Supply|Electricity                          Mt CO2/yr          
 73 Carbon Sequestration|CCS|Fossil|Energy|Supply|Liquids                              Mt CO2/yr          
 74 Carbon Sequestration|Land Use                                                      Mt CO2/yr          
 75 Concentration|CH4                                                                  ppb                
 76 Concentration|CO2                                                                  ppm                
 77 Concentration|N2O                                                                  ppb                
 78 Consumption                                                                        billion US$2010/yr 
 79 Diagnostics|MAGICC6|Harmonized Input|Emissions|BC                                  Mt BC/yr           
 80 Diagnostics|MAGICC6|Harmonized Input|Emissions|C2F6                                kt C2F6/yr         
 81 Diagnostics|MAGICC6|Harmonized Input|Emissions|C6F14                               kt C6F14/yr        
 82 Diagnostics|MAGICC6|Harmonized Input|Emissions|CH4                                 Mt CH4/yr          
 83 Diagnostics|MAGICC6|Harmonized Input|Emissions|CO                                  Mt CO/yr           
 84 Diagnostics|MAGICC6|Harmonized Input|Emissions|CO2|AFOLU                           Mt CO2/yr          
 85 Diagnostics|MAGICC6|Harmonized Input|Emissions|CO2|Energy and Industrial Processes Mt CO2/yr          
 86 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC125                          kt HFC23/yr        
 87 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC134a                         kt HFC23/yr        
 88 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC143a                         kt HFC23/yr        
 89 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC227ea                        kt HFC23/yr        
 90 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC23                           kt HFC23/yr        
 91 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC245ca                        kt HFC23/yr        
 92 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC32                           kt HFC23/yr        
 93 Diagnostics|MAGICC6|Harmonized Input|Emissions|HFC|HFC43-10                        kt HFC23/yr        
 94 Diagnostics|MAGICC6|Harmonized Input|Emissions|N2O                                 kt N2O/yr          
 95 Diagnostics|MAGICC6|Harmonized Input|Emissions|NH3                                 Mt NH3/yr          
 96 Diagnostics|MAGICC6|Harmonized Input|Emissions|NOx                                 Mt NO2/yr          
 97 Diagnostics|MAGICC6|Harmonized Input|Emissions|OC                                  Mt OC/yr           
 98 Diagnostics|MAGICC6|Harmonized Input|Emissions|PFC                                 kt CF4/yr          
 99 Diagnostics|MAGICC6|Harmonized Input|Emissions|SF6                                 kt SF6/yr          
100 Diagnostics|MAGICC6|Harmonized Input|Emissions|Sulfur                              Mt SO2/yr          
101 Diagnostics|MAGICC6|Harmonized Input|Emissions|VOC                                 Mt VOC/yr          
102 Emissions|BC                                                                       Mt BC/yr           
103 Emissions|CH4                                                                      Mt CH4/yr          
104 Emissions|CH4|AFOLU                                                                Mt CH4/yr          
105 Emissions|CH4|Energy|Demand|Industry                                               Mt CH4/yr          
106 Emissions|CH4|Energy|Demand|Residential and Commercial                             Mt CH4/yr          
107 Emissions|CH4|Energy|Demand|Transportation                                         Mt CH4/yr          
108 Emissions|CH4|Energy|Supply                                                        Mt CH4/yr          
109 Emissions|CO                                                                       Mt CO/yr           
110 Emissions|CO2                                                                      Mt CO2/yr          
111 Emissions|CO2|AFOLU                                                                Mt CO2/yr          
112 Emissions|CO2|Energy and Industrial Processes                                      Mt CO2/yr          
113 Emissions|CO2|Energy|Demand                                                        Mt CO2/yr          
114 Emissions|CO2|Energy|Demand|AFOFI                                                  Mt CO2/yr          
115 Emissions|CO2|Energy|Demand|Industry                                               Mt CO2/yr          
116 Emissions|CO2|Energy|Demand|Residential and Commercial                             Mt CO2/yr          
117 Emissions|CO2|Energy|Demand|Transportation                                         Mt CO2/yr          
118 Emissions|CO2|Energy|Supply                                                        Mt CO2/yr          
119 Emissions|CO2|Other                                                                Mt CO2/yr          
120 Emissions|F-Gases                                                                  Mt CO2-equiv/yr    
121 Emissions|HFC                                                                      kt HFC134a-equiv/yr
122 Emissions|HFC|HFC125                                                               kt HFC125/yr       
123 Emissions|HFC|HFC134a                                                              kt HFC134a/yr      
124 Emissions|HFC|HFC143a                                                              kt HFC143a/yr      
125 Emissions|HFC|HFC227ea                                                             kt HFC227ea/yr     
126 Emissions|HFC|HFC23                                                                kt HFC23/yr        
127 Emissions|HFC|HFC32                                                                kt HFC32/yr        
128 Emissions|HFC|HFC43-10                                                             kt HFC43-10/yr     
129 Emissions|Kyoto Gases (AR4-GWP100)                                                 Mt CO2-equiv/yr    
130 Emissions|Kyoto Gases (AR5-GWP100)                                                 Mt CO2-equiv/yr    
131 Emissions|Kyoto Gases (SAR-GWP100)                                                 Mt CO2-equiv/yr    
132 Emissions|N2O                                                                      kt N2O/yr          
133 Emissions|N2O|AFOLU                                                                kt N2O/yr          
134 Emissions|NH3                                                                      Mt NH3/yr          
135 Emissions|NOx                                                                      Mt NO2/yr          
136 Emissions|OC                                                                       Mt OC/yr           
137 Emissions|PFC                                                                      kt CF4-equiv/yr    
138 Emissions|SF6                                                                      kt SF6/yr          
139 Emissions|Sulfur                                                                   Mt SO2/yr          
140 Emissions|VOC                                                                      Mt VOC/yr          
141 Final Energy                                                                       EJ/yr              
142 Final Energy|Electricity                                                           EJ/yr              
143 Final Energy|Gases                                                                 EJ/yr              
144 Final Energy|Heat                                                                  EJ/yr              
145 Final Energy|Industry|Electricity                                                  EJ/yr              
146 Final Energy|Industry|Gases                                                        EJ/yr              
147 Final Energy|Industry|Heat                                                         EJ/yr              
148 Final Energy|Industry|Liquids                                                      EJ/yr              
149 Final Energy|Industry|Solids                                                       EJ/yr              
150 Final Energy|Liquids                                                               EJ/yr              
151 Final Energy|Residential and Commercial                                            EJ/yr              
152 Final Energy|Residential and Commercial|Heat                                       EJ/yr              
153 Final Energy|Solids                                                                EJ/yr              
154 Final Energy|Solids|Biomass                                                        EJ/yr              
155 Final Energy|Solids|Coal                                                           EJ/yr              
156 Final Energy|Transportation                                                        EJ/yr              
157 Final Energy|Transportation|Electricity                                            EJ/yr              
158 Final Energy|Transportation|Liquids|Biomass                                        EJ/yr              
159 Final Energy|Transportation|Liquids|Oil                                            EJ/yr              
160 Forcing                                                                            W/m2               
161 Forcing|Aerosol                                                                    W/m2               
162 Forcing|CH4                                                                        W/m2               
163 Forcing|CO2                                                                        W/m2               
164 Forcing|F-Gases                                                                    W/m2               
165 Forcing|Kyoto Gases                                                                W/m2               
166 Forcing|N2O                                                                        W/m2               
167 Forcing|Tropospheric Ozone                                                         W/m2               
168 GDP|MER                                                                            billion US$2010/yr 
169 GDP|PPP                                                                            billion US$2010/yr 
170 Investment|Energy Supply                                                           billion US$2010/yr 
171 Investment|Energy Supply|Electricity                                               billion US$2010/yr 
172 Investment|Energy Supply|Electricity|Fossil                                        billion US$2010/yr 
173 Investment|Energy Supply|Electricity|Non-fossil                                    billion US$2010/yr 
174 Investment|Energy Supply|Extraction|Fossil                                         billion US$2010/yr 
175 Land Cover                                                                         million ha         
176 Land Cover|Cropland                                                                million ha         
177 Land Cover|Cropland|Energy Crops                                                   million ha         
178 Land Cover|Forest                                                                  million ha         
179 Land Cover|Other Arable Land                                                       million ha         
180 Land Cover|Pasture                                                                 million ha         
181 Policy Cost|Consumption Loss                                                       billion US$2010/yr 
182 Policy Cost|GDP Loss                                                               billion US$2010/yr 
183 Population                                                                         million            
184 Price|Carbon                                                                       US$2010/t CO2      
185 Price|Primary Energy|Coal                                                          US$2010/GJ         
186 Price|Primary Energy|Gas                                                           US$2010/GJ         
187 Price|Primary Energy|Oil                                                           US$2010/GJ         
188 Price|Secondary Energy|Electricity                                                 US$2010/GJ         
189 Primary Energy                                                                     EJ/yr              
190 Primary Energy|Biomass                                                             EJ/yr              
191 Primary Energy|Coal                                                                EJ/yr              
192 Primary Energy|Coal|w/ CCS                                                         EJ/yr              
193 Primary Energy|Coal|w/o CCS                                                        EJ/yr              
194 Primary Energy|Fossil                                                              EJ/yr              
195 Primary Energy|Fossil|w/ CCS                                                       EJ/yr              
196 Primary Energy|Fossil|w/o CCS                                                      EJ/yr              
197 Primary Energy|Gas                                                                 EJ/yr              
198 Primary Energy|Gas|w/ CCS                                                          EJ/yr              
199 Primary Energy|Gas|w/o CCS                                                         EJ/yr              
200 Primary Energy|Non-Biomass Renewables                                              EJ/yr              
201 Primary Energy|Nuclear                                                             EJ/yr              
202 Primary Energy|Oil                                                                 EJ/yr              
203 Primary Energy|Oil|w/ CCS                                                          EJ/yr              
204 Primary Energy|Oil|w/o CCS                                                         EJ/yr              
205 Primary Energy|Other                                                               EJ/yr              
206 Primary Energy|Secondary Energy Trade                                              EJ/yr              
207 Secondary Energy|Electricity                                                       EJ/yr              
208 Secondary Energy|Electricity|Biomass                                               EJ/yr              
209 Secondary Energy|Electricity|Biomass|w/ CCS                                        EJ/yr              
210 Secondary Energy|Electricity|Biomass|w/o CCS                                       EJ/yr              
211 Secondary Energy|Electricity|Coal                                                  EJ/yr              
212 Secondary Energy|Electricity|Coal|w/ CCS                                           EJ/yr              
213 Secondary Energy|Electricity|Coal|w/o CCS                                          EJ/yr              
214 Secondary Energy|Electricity|Gas                                                   EJ/yr              
215 Secondary Energy|Electricity|Gas|w/ CCS                                            EJ/yr              
216 Secondary Energy|Electricity|Gas|w/o CCS                                           EJ/yr              
217 Secondary Energy|Electricity|Geothermal                                            EJ/yr              
218 Secondary Energy|Electricity|Hydro                                                 EJ/yr              
219 Secondary Energy|Electricity|Non-Biomass Renewables                                EJ/yr              
220 Secondary Energy|Electricity|Nuclear                                               EJ/yr              
221 Secondary Energy|Electricity|Oil                                                   EJ/yr              
222 Secondary Energy|Electricity|Oil|w/ CCS                                            EJ/yr              
223 Secondary Energy|Electricity|Oil|w/o CCS                                           EJ/yr              
224 Secondary Energy|Electricity|Solar                                                 EJ/yr              
225 Secondary Energy|Electricity|Wind                                                  EJ/yr              
226 Secondary Energy|Electricity|Wind|Onshore                                          EJ/yr              
227 Secondary Energy|Gases                                                             EJ/yr              
228 Secondary Energy|Gases|Natural Gas                                                 EJ/yr              
229 Secondary Energy|Heat                                                              EJ/yr              
230 Secondary Energy|Liquids                                                           EJ/yr              
231 Secondary Energy|Liquids|Biomass                                                   EJ/yr              
232 Secondary Energy|Liquids|Oil                                                       EJ/yr              
233 Temperature|Global Mean                                                            °C                 
234 Agricultural Demand                                                                million t DM/yr    
235 Agricultural Demand|Crops|Feed                                                     million t DM/yr    
236 Agricultural Demand|Crops|Food                                                     million t DM/yr    
237 Agricultural Demand|Crops|Other                                                    million t DM/yr    
238 Agricultural Demand|Livestock|Food                                                 million t DM/yr    
239 Agricultural Demand|Livestock|Other                                                million t DM/yr    
240 Agricultural Production                                                            million t DM/yr    
241 Capacity|Electricity                                                               GW                 
242 Capacity|Electricity|Oil                                                           GW                 
243 Cumulative Capacity|Electricity|Biomass                                            GW                 
244 Cumulative Capacity|Electricity|Biomass|w/ CCS                                     GW                 
245 Cumulative Capacity|Electricity|Coal                                               GW                 
246 Cumulative Capacity|Electricity|Coal|w/ CCS                                        GW                 
247 Cumulative Capacity|Electricity|Coal|w/o CCS                                       GW                 
248 Cumulative Capacity|Electricity|Gas                                                GW                 
249 Cumulative Capacity|Electricity|Gas|w/ CCS                                         GW                 
250 Cumulative Capacity|Electricity|Gas|w/o CCS                                        GW                 
251 Cumulative Capacity|Electricity|Nuclear                                            GW                 
252 Cumulative Capacity|Electricity|Oil                                                GW                 
253 Cumulative Capacity|Electricity|Solar                                              GW                 
254 Cumulative Capacity|Electricity|Solar|PV                                           GW                 
255 Emissions|BC|AFOLU                                                                 Mt BC/yr           
256 Emissions|BC|Energy                                                                Mt BC/yr           
257 Emissions|BC|Energy|Demand|Industry                                                Mt BC/yr           
258 Emissions|BC|Energy|Demand|Residential and Commercial                              Mt BC/yr           
259 Emissions|BC|Energy|Demand|Transportation                                          Mt BC/yr           
260 Emissions|BC|Energy|Supply                                                         Mt BC/yr           
261 Emissions|CH4|Energy                                                               Mt CH4/yr          
262 Emissions|CO2|Energy                                                               Mt CO2/yr          
263 Emissions|CO2|Energy|Supply|Other Sector                                           Mt CO2/yr          
264 Emissions|CO2|Industrial Processes                                                 Mt CO2/yr          
265 Emissions|CO|AFOLU                                                                 Mt CO/yr           
266 Emissions|CO|Energy                                                                Mt CO/yr           
267 Emissions|CO|Energy|Demand|Industry                                                Mt CO/yr           
268 Emissions|CO|Energy|Demand|Residential and Commercial                              Mt CO/yr           
269 Emissions|CO|Energy|Demand|Transportation                                          Mt CO/yr           
270 Emissions|CO|Energy|Supply                                                         Mt CO/yr           
271 Emissions|Kyoto Gases                                                              Mt CO2-equiv/yr    
272 Emissions|N2O|Energy                                                               kt N2O/yr          
273 Emissions|N2O|Other                                                                kt N2O/yr          
274 Emissions|NH3|AFOLU                                                                Mt NH3/yr          
275 Emissions|NH3|Energy                                                               Mt NH3/yr          
276 Emissions|NH3|Energy|Demand|Industry                                               Mt NH3/yr          
277 Emissions|NH3|Energy|Demand|Residential and Commercial                             Mt NH3/yr          
278 Emissions|NH3|Energy|Demand|Transportation                                         Mt NH3/yr          
279 Emissions|NH3|Energy|Supply                                                        Mt NH3/yr          
280 Emissions|NOx|AFOLU                                                                Mt NO2/yr          
281 Emissions|NOx|Energy                                                               Mt NO2/yr          
282 Emissions|NOx|Energy|Demand|Industry                                               Mt NO2/yr          
283 Emissions|NOx|Energy|Demand|Residential and Commercial                             Mt NO2/yr          
284 Emissions|NOx|Energy|Demand|Transportation                                         Mt NO2/yr          
285 Emissions|NOx|Energy|Supply                                                        Mt NO2/yr          
286 Emissions|OC|AFOLU                                                                 Mt OC/yr           
287 Emissions|OC|Energy                                                                Mt OC/yr           
288 Emissions|OC|Energy|Demand|Industry                                                Mt OC/yr           
289 Emissions|OC|Energy|Demand|Residential and Commercial                              Mt OC/yr           
290 Emissions|OC|Energy|Demand|Transportation                                          Mt OC/yr           
291 Emissions|OC|Energy|Supply                                                         Mt OC/yr           
292 Emissions|Sulfur|AFOLU                                                             Mt SO2/yr          
293 Emissions|Sulfur|Energy                                                            Mt SO2/yr          
294 Emissions|Sulfur|Energy|Demand|Industry                                            Mt SO2/yr          
295 Emissions|Sulfur|Energy|Demand|Residential and Commercial                          Mt SO2/yr          
296 Emissions|Sulfur|Energy|Demand|Transportation                                      Mt SO2/yr          
297 Emissions|Sulfur|Energy|Supply                                                     Mt SO2/yr          
298 Emissions|VOC|AFOLU                                                                Mt VOC/yr          
299 Emissions|VOC|Energy                                                               Mt VOC/yr          
300 Emissions|VOC|Energy|Demand|Industry                                               Mt VOC/yr          
301 Emissions|VOC|Energy|Demand|Residential and Commercial                             Mt VOC/yr          
302 Emissions|VOC|Energy|Demand|Transportation                                         Mt VOC/yr          
303 Emissions|VOC|Energy|Supply                                                        Mt VOC/yr          
304 Final Energy|Industry                                                              EJ/yr              
305 Final Energy|Industry|Solids|Biomass                                               EJ/yr              
306 Final Energy|Industry|Solids|Coal                                                  EJ/yr              
307 Final Energy|Residential and Commercial|Electricity                                EJ/yr              
308 Final Energy|Residential and Commercial|Gases                                      EJ/yr              
309 Final Energy|Residential and Commercial|Liquids                                    EJ/yr              
310 Final Energy|Residential and Commercial|Solids                                     EJ/yr              
311 Final Energy|Residential and Commercial|Solids|Biomass                             EJ/yr              
312 Final Energy|Residential and Commercial|Solids|Coal                                EJ/yr              
313 Final Energy|Transportation|Gases                                                  EJ/yr              
314 Final Energy|Transportation|Liquids                                                EJ/yr              
315 Food Energy Supply                                                                 EJ/yr              
316 Forcing|Aerosol|BC                                                                 W/m2               
317 Forcing|Aerosol|OC                                                                 W/m2               
318 Forcing|Aerosol|Other                                                              W/m2               
319 Forcing|Aerosol|Sulfate Direct                                                     W/m2               
320 Forcing|Other                                                                      W/m2               
321 Investment|Energy Supply|CO2 Transport and Storage                                 billion US$2010/yr 
322 Investment|Energy Supply|Electricity|Other                                         billion US$2010/yr 
323 Investment|Energy Supply|Electricity|Transmission and Distribution                 billion US$2010/yr 
324 Investment|Energy Supply|Heat                                                      billion US$2010/yr 
325 Investment|Energy Supply|Liquids                                                   billion US$2010/yr 
326 Land Cover|Built-up Area                                                           million ha         
327 Land Cover|Cropland|Cereals                                                        million ha         
328 Land Cover|Forest|Natural Forest                                                   million ha         
329 Land Cover|Other Land                                                              million ha         
330 Price|Agriculture|Non-Energy Crops and Livestock|Index                             Index (2010 = 1)   
331 Price|Agriculture|Non-Energy Crops|Index                                           Index (2010 = 1)   
332 Price|Secondary Energy|Liquids|Biomass                                             US$2010/GJ         
333 Secondary Energy|Heat|Geothermal                                                   EJ/yr              
334 Trade|Primary Energy|Coal|Volume                                                   EJ/yr              
335 Trade|Primary Energy|Gas|Volume                                                    EJ/yr              
336 Trade|Primary Energy|Oil|Volume                                                    EJ/yr              
337 Agricultural Demand|Crops                                                          million t DM/yr    
338 Agricultural Demand|Livestock                                                      million t DM/yr    
339 Agricultural Production|Energy|Crops                                               million t DM/yr    
340 Agricultural Production|Non-Energy|Crops                                           million t DM/yr    
341 Agricultural Production|Non-Energy|Livestock                                       million t DM/yr    
342 Emissions|CO2|Energy|Supply|Electricity                                            Mt CO2/yr          
343 Food Demand                                                                        kcal/cap/day       
344 Food Demand|Crops                                                                  kcal/cap/day       
345 Food Demand|Livestock                                                              kcal/cap/day       
346 Land Cover|Other Natural Land                                                      million ha         
347 Price|Agriculture|Livestock|Index                                                  Index (2010 = 1)   
348 Primary Energy|Biomass|Modern|w/ CCS                                               EJ/yr              
349 Primary Energy|Biomass|Modern|w/o CCS                                              EJ/yr              
350 Primary Energy|Geothermal                                                          EJ/yr              
351 Primary Energy|Hydro                                                               EJ/yr              
352 Primary Energy|Solar                                                               EJ/yr              
353 Primary Energy|Wind                                                                EJ/yr              
354 Carbon Sequestration|Land Use|Afforestation                                        Mt CO2/yr          
355 Final Energy|Geothermal                                                            EJ/yr              
356 Final Energy|Hydrogen                                                              EJ/yr              
357 Final Energy|Solar                                                                 EJ/yr              
358 Investment|Energy Efficiency                                                       billion US$2010/yr 
359 Investment|Energy Supply|Electricity|Biomass|w/ CCS                                billion US$2010/yr 
360 Investment|Energy Supply|Electricity|Biomass|w/o CCS                               billion US$2010/yr 
361 Investment|Energy Supply|Electricity|Coal|w/ CCS                                   billion US$2010/yr 
362 Investment|Energy Supply|Electricity|Coal|w/o CCS                                  billion US$2010/yr 
363 Investment|Energy Supply|Electricity|Gas|w/ CCS                                    billion US$2010/yr 
364 Investment|Energy Supply|Electricity|Gas|w/o CCS                                   billion US$2010/yr 
365 Investment|Energy Supply|Electricity|Geothermal                                    billion US$2010/yr 
366 Investment|Energy Supply|Electricity|Hydro                                         billion US$2010/yr 
367 Investment|Energy Supply|Electricity|Non-Biomass Renewables                        billion US$2010/yr 
368 Investment|Energy Supply|Electricity|Nuclear                                       billion US$2010/yr 
369 Investment|Energy Supply|Electricity|Oil|w/o CCS                                   billion US$2010/yr 
370 Investment|Energy Supply|Electricity|Solar                                         billion US$2010/yr 
371 Investment|Energy Supply|Electricity|Wind                                          billion US$2010/yr 
372 Investment|Energy Supply|Hydrogen|Fossil                                           billion US$2010/yr 
373 Investment|Energy Supply|Hydrogen|Other                                            billion US$2010/yr 
374 Price|Agriculture|Corn|Index                                                       Index (2010 = 1)   
375 Price|Agriculture|Soybean|Index                                                    Index (2010 = 1)   
376 Price|Agriculture|Wheat|Index                                                      Index (2010 = 1)   
377 Final Energy|Solids|Biomass|Traditional                                            EJ/yr              
378 Final Energy|Transportation|Freight                                                EJ/yr              
379 Primary Energy|Biomass|Electricity                                                 EJ/yr              
380 Primary Energy|Biomass|Electricity|w/o CCS                                         EJ/yr              
381 Primary Energy|Biomass|Modern                                                      EJ/yr              
382 Secondary Energy|Liquids|Biomass|w/o CCS                                           EJ/yr              
383 Secondary Energy|Solids                                                            EJ/yr              
384 Primary Energy|Biomass|Electricity|w/ CCS                                          EJ/yr              
385 Secondary Energy|Liquids|Biomass|w/ CCS                                            EJ/yr              
386 Capacity|Electricity|Other                                                         GW                 
387 Capital Cost|Electricity|Biomass|w/ CCS                                            US$2010/kW         
388 Capital Cost|Electricity|Biomass|w/o CCS                                           US$2010/kW         
389 Capital Cost|Electricity|Coal|w/ CCS                                               US$2010/kW         
390 Capital Cost|Electricity|Coal|w/o CCS                                              US$2010/kW         
391 Capital Cost|Electricity|Gas|w/ CCS                                                US$2010/kW         
392 Capital Cost|Electricity|Gas|w/o CCS                                               US$2010/kW         
393 Capital Cost|Electricity|Geothermal                                                US$2010/kW         
394 Capital Cost|Electricity|Hydro                                                     US$2010/kW         
395 Capital Cost|Electricity|Nuclear                                                   US$2010/kW         
396 Capital Cost|Electricity|Solar|PV                                                  US$2010/kW         
397 Capital Cost|Electricity|Wind|Onshore                                              US$2010/kW         
398 Carbon Sequestration|CCS|Biomass|Energy                                            Mt CO2/yr          
399 Carbon Sequestration|CCS|Fossil|Energy                                             Mt CO2/yr          
400 Carbon Sequestration|CCS|Fossil|Industrial Processes                               Mt CO2/yr          
401 Cumulative Capacity|Electricity|Biomass|w/o CCS                                    GW                 
402 Cumulative Capacity|Electricity|Geothermal                                         GW                 
403 Cumulative Capacity|Electricity|Hydro                                              GW                 
404 Cumulative Capacity|Electricity|Other                                              GW                 
405 Cumulative Capacity|Electricity|Wind|Onshore                                       GW                 
406 Emissions|HFC|HFC245fa                                                             kt HFC245fa/yr     
407 Final Energy|Non-Energy Use                                                        EJ/yr              
408 Final Energy|Non-Energy Use|Gas                                                    EJ/yr              
409 Final Energy|Non-Energy Use|Oil                                                    EJ/yr              
410 Food Energy Supply|Livestock                                                       EJ/yr              
411 Forcing|Aerosol|Black and Organic Carbon                                           W/m2               
412 Forcing|Aerosol|Cloud Indirect                                                     W/m2               
413 Investment|Energy Supply|Electricity|Biomass                                       billion US$2010/yr 
414 Investment|Energy Supply|Electricity|Coal                                          billion US$2010/yr 
415 Investment|Energy Supply|Electricity|Electricity Storage                           billion US$2010/yr 
416 Investment|Energy Supply|Electricity|Gas                                           billion US$2010/yr 
417 Investment|Energy Supply|Electricity|Oil                                           billion US$2010/yr 
418 Investment|Energy Supply|Electricity|Oil|w/ CCS                                    billion US$2010/yr 
419 Investment|Energy Supply|Extraction|Coal                                           billion US$2010/yr 
420 Investment|Energy Supply|Extraction|Gas                                            billion US$2010/yr 
421 Investment|Energy Supply|Extraction|Oil                                            billion US$2010/yr 
422 Investment|Energy Supply|Liquids|Biomass                                           billion US$2010/yr 
423 Investment|Energy Supply|Liquids|Oil                                               billion US$2010/yr 
424 Population|Risk of Hunger                                                          Million            
425 Price|Final Energy|Residential|Electricity                                         US$2010/GJ         
426 Price|Final Energy|Residential|Gases|Natural Gas                                   US$2010/GJ         
427 Price|Final Energy|Residential|Liquids|Biomass                                     US$2010/GJ         
428 Price|Final Energy|Residential|Liquids|Oil                                         US$2010/GJ         
429 Price|Final Energy|Residential|Solids|Coal                                         US$2010/GJ         
430 Price|Secondary Energy|Liquids                                                     US$2010/GJ         
431 Primary Energy|Biomass|Traditional                                                 EJ/yr              
432 Primary Energy|Coal|Electricity                                                    EJ/yr              
433 Primary Energy|Coal|Electricity|w/ CCS                                             EJ/yr              
434 Primary Energy|Coal|Electricity|w/o CCS                                            EJ/yr              
435 Primary Energy|Gas|Electricity                                                     EJ/yr              
436 Primary Energy|Gas|Electricity|w/ CCS                                              EJ/yr              
437 Primary Energy|Gas|Electricity|w/o CCS                                             EJ/yr              
438 Secondary Energy|Electricity|Solar|PV                                              EJ/yr              
439 Trade|Primary Energy|Biomass|Volume                                                EJ/yr              
440 Trade|Secondary Energy|Electricity|Volume                                          EJ/yr              
441 Trade|Secondary Energy|Liquids|Biomass|Volume                                      EJ/yr              
442 Water Consumption|Electricity                                                      km3/yr             
443 Water Consumption|Industrial Water                                                 km3/yr             
444 Water Withdrawal|Electricity                                                       km3/yr             
445 Water Withdrawal|Industrial Water                                                  km3/yr             
446 Carbon Sequestration|Direct Air Capture                                            Mt CO2/yr          
447 Carbon Sequestration|Enhanced Weathering                                           Mt CO2/yr          
448 Carbon Sequestration|Land Use|Biochar                                              Mt CO2/yr          
449 Carbon Sequestration|Land Use|Soil Carbon Management                               Mt CO2/yr          
450 Forcing|Montreal Gases                                                             W/m2               
451 Final Energy|Transportation|Hydrogen                                               EJ/yr              
452 Secondary Energy|Gases|Biomass                                                     EJ/yr              
453 Secondary Energy|Gases|Coal                                                        EJ/yr              
454 Secondary Energy|Hydrogen                                                          EJ/yr              
455 Secondary Energy|Hydrogen|Biomass                                                  EJ/yr              
456 Secondary Energy|Hydrogen|Biomass|w/ CCS                                           EJ/yr              
457 Secondary Energy|Hydrogen|Biomass|w/o CCS                                          EJ/yr              
458 Secondary Energy|Hydrogen|Electricity                                              EJ/yr              
459 Secondary Energy|Liquids|Coal                                                      EJ/yr              
460 Secondary Energy|Liquids|Coal|w/ CCS                                               EJ/yr              
461 Secondary Energy|Liquids|Coal|w/o CCS                                              EJ/yr              
462 Secondary Energy|Liquids|Gas                                                       EJ/yr              
463 Secondary Energy|Liquids|Gas|w/ CCS                                                EJ/yr              
464 Secondary Energy|Liquids|Gas|w/o CCS                                               EJ/yr              
465 Capacity|Electricity|Ocean                                                         GW                 
466 Capacity|Electricity|Solar|CSP                                                     GW                 
467 Capacity|Electricity|Wind|Offshore                                                 GW                 
468 Capital Cost|Electricity|Solar|CSP                                                 US$2010/kW         
469 Capital Cost|Electricity|Wind|Offshore                                             US$2010/kW         
470 Cumulative Capacity|Electricity|Ocean                                              GW                 
471 Cumulative Capacity|Electricity|Solar|CSP                                          GW                 
472 Cumulative Capacity|Electricity|Wind                                               GW                 
473 Cumulative Capacity|Electricity|Wind|Offshore                                      GW                 
474 Discount rate|Buildings                                                            %                  
475 Discount rate|Economy                                                              %                  
476 Discount rate|Electricity                                                          %                  
477 Discount rate|Energy Supply                                                        %                  
478 Discount rate|Industry                                                             %                  
479 Discount rate|Mobility                                                             %                  
480 Discount rate|Social                                                               %                  
481 Discount rate|Transport Infrastructure                                             %                  
482 Energy Service|Transportation|Freight                                              bn tkm/yr          
483 Energy Service|Transportation|Freight|Navigation                                   bn tkm/yr          
484 Energy Service|Transportation|Freight|Railways                                     bn tkm/yr          
485 Energy Service|Transportation|Freight|Road                                         bn tkm/yr          
486 Energy Service|Transportation|Passenger                                            bn pkm/yr          
487 Energy Service|Transportation|Passenger|Aviation                                   bn pkm/yr          
488 Energy Service|Transportation|Passenger|Railways                                   bn pkm/yr          
489 Energy Service|Transportation|Passenger|Road                                       bn pkm/yr          
490 Primary Energy|Ocean                                                               EJ/yr              
491 Secondary Energy|Electricity|Ocean                                                 EJ/yr              
492 Secondary Energy|Electricity|Solar|CSP                                             EJ/yr              
493 Secondary Energy|Electricity|Wind|Offshore                                         EJ/yr              
494 Final Energy|Industry|Hydrogen                                                     EJ/yr              
495 Final Energy|Industry|Other                                                        EJ/yr              
496 Final Energy|Non-Energy Use|Biomass                                                EJ/yr              
497 Final Energy|Non-Energy Use|Coal                                                   EJ/yr              
498 Final Energy|Residential and Commercial|Hydrogen                                   EJ/yr              
499 Final Energy|Residential and Commercial|Other                                      EJ/yr              
500 Population|Rural                                                                   million            
501 Population|Urban                                                                   million            
502 Capital Cost|Hydrogen|Biomass|w/ CCS                                               US$2010/kW         
503 Capital Cost|Hydrogen|Biomass|w/o CCS                                              US$2010/kW         
504 Capital Cost|Hydrogen|Coal|w/ CCS                                                  US$2010/kW         
505 Capital Cost|Hydrogen|Coal|w/o CCS                                                 US$2010/kW         
506 Capital Cost|Hydrogen|Electricity                                                  US$2010/kW         
507 Capital Cost|Hydrogen|Gas|w/ CCS                                                   US$2010/kW         
508 Capital Cost|Hydrogen|Gas|w/o CCS                                                  US$2010/kW         
509 Carbon Sequestration|CCS|Biomass|Energy|Demand|Industry                            Mt CO2/yr          
510 Emissions|CH4|Other                                                                Mt CH4/yr          
511 Emissions|CO2|Energy|Demand|Other Sector                                           Mt CO2/yr          
512 Emissions|CO2|Energy|Supply|Heat                                                   Mt CO2/yr          
513 Emissions|CO2|Energy|Supply|Liquids                                                Mt CO2/yr          
514 Emissions|PFC|C2F6                                                                 kt C2F6/yr         
515 Emissions|PFC|C6F14                                                                kt C6F14/yr        
516 Emissions|PFC|CF4                                                                  kt CF4/yr          
517 Investment|Energy Supply|Electricity|Ocean                                         billion US$2010/yr 
518 Investment|Energy Supply|Liquids|Coal and Gas                                      billion US$2010/yr 
519 Population|Urban|Share                                                             %                  
520 Carbon Sequestration|CCS|Biomass|Energy|Supply|Hydrogen                            Mt CO2/yr          
521 Carbon Sequestration|CCS|Biomass|Energy|Supply|Other                               Mt CO2/yr          
522 Carbon Sequestration|CCS|Fossil|Energy|Supply|Gases                                Mt CO2/yr          
523 Carbon Sequestration|CCS|Fossil|Energy|Supply|Hydrogen                             Mt CO2/yr          
524 Carbon Sequestration|CCS|Fossil|Energy|Supply|Other                                Mt CO2/yr          
525 Emissions|CO2|Energy|Supply|Gases                                                  Mt CO2/yr          
526 Emissions|CO2|Energy|Supply|Solids                                                 Mt CO2/yr          
527 Forcing|Albedo Change and Mineral Dust                                             W/m2               
528 Land Cover|Forest|Managed                                                          million ha         
529 Policy Cost|Area under MAC Curve                                                   billion US$2010/yr 
530 Value Added|Commercial                                                             billion US$2010/yr 
531 Value Added|Industry                                                               billion US$2010/yr 
532 Agricultural Demand|Crops|Energy                                                   million t DM/yr    
533 Agricultural Production|Energy|Residues                                            million t DM/yr    
534 Agricultural Production|Non-Energy                                                 million t DM/yr    
535 Carbon Sequestration|CCS|Biomass|Industrial Processes                              Mt CO2/yr          
536 Energy Service|Transportation|Freight|Aviation                                     bn tkm/yr          
537 Land Cover|Cropland|Irrigated                                                      million ha         
538 Land Cover|Forest|Afforestation and Reforestation                                  million ha         
539 Price|Secondary Energy|Gases|Natural Gas                                           US$2010/GJ         
540 Price|Secondary Energy|Liquids|Oil                                                 US$2010/GJ         
541 Price|Secondary Energy|Solids|Biomass                                              US$2010/GJ         
542 Price|Secondary Energy|Solids|Coal                                                 US$2010/GJ         
543 Secondary Energy|Hydrogen|Fossil                                                   EJ/yr              
544 Secondary Energy|Hydrogen|Fossil|w/ CCS                                            EJ/yr              
545 Secondary Energy|Hydrogen|Fossil|w/o CCS                                           EJ/yr              
546 Price|Primary Energy|Biomass                                                       US$2010/GJ         
547 Price|Secondary Energy|Hydrogen                                                    US$2010/GJ         
548 Agricultural Production|Energy                                                     million t DM/yr    
549 Investment|Energy Supply|Hydrogen                                                  billion US$2010/yr 
550 Water Consumption                                                                  km3/yr             
551 Water Consumption|Extraction                                                       km3/yr             
552 Water Consumption|Irrigation                                                       km3/yr             
553 Water Withdrawal                                                                   km3/yr             
554 Water Withdrawal|Extraction                                                        km3/yr             
555 Water Withdrawal|Irrigation                                                        km3/yr             
556 Carbon Sequestration|CCS|Biomass|Energy|Supply|Gases                               Mt CO2/yr          
557 Policy Cost|Other                                                                  billion US$2010/yr 
558 Capital Cost|Gases|Biomass|w/o CCS                                                 US$2010/kW         
559 Capital Cost|Gases|Coal|w/o CCS                                                    US$2010/kW         
560 Capital Cost|Liquids|Biomass|w/ CCS                                                US$2010/kW         
561 Capital Cost|Liquids|Biomass|w/o CCS                                               US$2010/kW         
562 Capital Cost|Liquids|Coal|w/ CCS                                                   US$2010/kW         
563 Capital Cost|Liquids|Coal|w/o CCS                                                  US$2010/kW         
564 Capital Cost|Liquids|Oil                                                           US$2010/kW         
565 Emissions|BC|Other                                                                 Mt BC/yr           
566 Emissions|NH3|Other                                                                Mt NH3/yr          
567 Emissions|NOx|Other                                                                Mt NO2/yr          
568 Emissions|OC|Other                                                                 Mt OC/yr           
569 Emissions|Sulfur|Other                                                             Mt SO2/yr          
570 Emissions|VOC|Other                                                                Mt VOC/yr          
571 Price|Final Energy|Residential|Solids|Biomass                                      US$2010/GJ         
572 Subsidies|Energy|Fossil                                                            billion US$2010/yr 
573 Emissions|CO|Other                                                                 Mt CO/yr           
574 Final Energy|Fossil                                                                EJ/yr              
575 Final Energy|Industry|Fossil                                                       EJ/yr              
576 Final Energy|Industry|Geothermal                                                   EJ/yr              
577 Final Energy|Industry|Solar                                                        EJ/yr              
578 Final Energy|Non-Energy Use|Fossil                                                 EJ/yr              
579 Final Energy|Residential and Commercial|Fossil                                     EJ/yr              
580 Final Energy|Residential and Commercial|Geothermal                                 EJ/yr              
581 Final Energy|Residential and Commercial|Solar                                      EJ/yr              
582 Final Energy|Transportation|Fossil                                                 EJ/yr              
583 Final Energy|Transportation|Geothermal                                             EJ/yr              
584 Final Energy|Transportation|Heat                                                   EJ/yr              
585 Final Energy|Transportation|Solar                                                  EJ/yr              
586 Final Energy|Transportation|Solids|Biomass                                         EJ/yr              
587 Final Energy|Transportation|Solids|Coal                                            EJ/yr              
588 Secondary Energy|Electricity|Fossil                                                EJ/yr              
589 Secondary Energy|Electricity|Other                                                 EJ/yr              
590 Policy Cost|Additional Total Energy System Cost                                    billion US$2010/yr
```

#### Sceanrios

```
  1 ADVANCE_2020_1.5C-2100                
  2 ADVANCE_2020_Med2C                    
  3 ADVANCE_2020_WB2C                     
  4 ADVANCE_2030_Med2C                    
  5 ADVANCE_2030_Price1.5C                
  6 ADVANCE_2030_WB2C                     
  7 ADVANCE_INDC                          
  8 ADVANCE_NoPolicy                      
  9 ADVANCE_Reference                     
 10 SFCM_SSP2_Bio_1p5Degree               
 11 SFCM_SSP2_Bio_2Degree                 
 12 SFCM_SSP2_Bio_Baseline                
 13 SFCM_SSP2_EEEI_1p5Degree              
 14 SFCM_SSP2_EEEI_2Degree                
 15 SFCM_SSP2_EEEI_Baseline               
 16 SFCM_SSP2_LifeStyle_1p5Degree         
 17 SFCM_SSP2_LifeStyle_2Degree           
 18 SFCM_SSP2_LifeStyle_Baseline          
 19 SFCM_SSP2_Ref_1p5Degree               
 20 SFCM_SSP2_Ref_2Degree                 
 21 SFCM_SSP2_Ref_Baseline                
 22 SFCM_SSP2_ST_CCS_1p5Degree            
 23 SFCM_SSP2_ST_CCS_2Degree              
 24 SFCM_SSP2_ST_CCS_Baseline             
 25 SFCM_SSP2_ST_bio_1p5Degree            
 26 SFCM_SSP2_ST_bio_2Degree              
 27 SFCM_SSP2_ST_bio_Baseline             
 28 SFCM_SSP2_ST_nuclear_1p5Degree        
 29 SFCM_SSP2_ST_nuclear_2Degree          
 30 SFCM_SSP2_ST_nuclear_Baseline         
 31 SFCM_SSP2_ST_solar_1p5Degree          
 32 SFCM_SSP2_ST_solar_2Degree            
 33 SFCM_SSP2_ST_solar_Baseline           
 34 SFCM_SSP2_ST_wind_1p5Degree           
 35 SFCM_SSP2_ST_wind_2Degree             
 36 SFCM_SSP2_ST_wind_Baseline            
 37 SFCM_SSP2_SupTech_1p5Degree           
 38 SFCM_SSP2_SupTech_2Degree             
 39 SFCM_SSP2_SupTech_Baseline            
 40 SFCM_SSP2_combined_1p5Degree          
 41 SFCM_SSP2_combined_2Degree            
 42 SFCM_SSP2_combined_Baseline           
 43 SSP1-19                               
 44 SSP1-26                               
 45 SSP1-34                               
 46 SSP1-45                               
 47 SSP1-Baseline                         
 48 SSP2-19                               
 49 SSP2-26                               
 50 SSP2-34                               
 51 SSP2-45                               
 52 SSP2-60                               
 53 SSP2-Baseline                         
 54 SSP3-34                               
 55 SSP3-45                               
 56 SSP3-60                               
 57 SSP3-Baseline                         
 58 SSP4-26                               
 59 SSP4-34                               
 60 SSP4-45                               
 61 SSP4-Baseline                         
 62 SSP5-26                               
 63 SSP5-34                               
 64 SSP5-45                               
 65 SSP5-60                               
 66 SSP5-Baseline                         
 67 CD-LINKS_INDCi                        
 68 CD-LINKS_NPi                          
 69 CD-LINKS_NPi2020_1000                 
 70 CD-LINKS_NPi2020_1600                 
 71 CD-LINKS_NPi2020_400                  
 72 CD-LINKS_NoPolicy                     
 73 EMF33_Baseline                        
 74 EMF33_Med2C_cost100                   
 75 EMF33_Med2C_full                      
 76 EMF33_Med2C_nobeccs                   
 77 EMF33_Med2C_nofuel                    
 78 EMF33_Med2C_none                      
 79 EMF33_WB2C_cost100                    
 80 EMF33_WB2C_full                       
 81 EMF33_tax_hi_full                     
 82 EMF33_tax_hi_none                     
 83 EMF33_tax_lo_full                     
 84 EMF33_tax_lo_none                     
 85 TERL_15D_LowCarbonTransportPolicy     
 86 TERL_15D_NoTransportPolicy            
 87 TERL_2D_LowCarbonTransportPolicy      
 88 TERL_2D_NoTransportPolicy             
 89 TERL_Baseline_LowCarbonTransportPolicy
 90 TERL_Baseline_NoTransportPolicy       
 91 Ratchet-1.5-allCDR                    
 92 Ratchet-1.5-limCDR                    
 93 Ratchet-1.5-limCDR-noOS               
 94 Ratchet-1.5-noCDR                     
 95 Ratchet-1.5-noCDR-noOS                
 96 Reference                             
 97 SSP4-60                               
 98 SSP5-19                               
 99 1.0                                   
100 B2DS                                  
101 Faster Transition Scenario            
102 IMA15-AGInt                           
103 IMA15-Def                             
104 IMA15-Eff                             
105 IMA15-LiStCh                          
106 IMA15-LoNCO2                          
107 IMA15-Pop                             
108 IMA15-RenElec                         
109 IMA15-TOT                             
110 EMF33_Med2C_limbio                    
111 EMF33_WB2C_limbio                     
112 EMF33_WB2C_nobeccs                    
113 EMF33_WB2C_nofuel                     
114 BAU                                   
115 DAC15_50                              
116 DAC2_66                               
117 GEA_Eff_1p5C                          
118 GEA_Eff_1p5C_Delay2020                
119 GEA_Eff_2C_Delay2020                  
120 GEA_Eff_AdvNCO2_1p5C                  
121 GEA_Eff_base                          
122 GEA_Mix_1p5C_AdvNCO2_PartialDelay2020 
123 GEA_Mix_1p5C_AdvTrans_PartialDelay2020
124 GEA_Mix_2C_AdvNCO2_PartialDelay2020   
125 GEA_Mix_2C_AdvTrans_PartialDelay2020  
126 GEA_Mix_base                          
127 EMF33_1.5C_cost100                    
128 EMF33_1.5C_full                       
129 LowEnergyDemand                       
130 ADVANCE_2030_1.5C-2100                
131 EMF33_1.5C_limbio                     
132 EMF33_1.5C_nofuel                     
133 EMF33_WB2C_none                       
134 EMC_Def_100$                          
135 EMC_Def_30$                           
136 EMC_LimSW_100$                        
137 EMC_LimSW_30$                         
138 EMC_NucPO_100$                        
139 EMC_NucPO_30$                         
140 EMC_lowEI_100$                        
141 EMC_lowEI_30$                         
142 CEMICS-1.5-CDR12                      
143 CEMICS-1.5-CDR20                      
144 CEMICS-1.5-CDR8                       
145 CEMICS-2.0-CDR12                      
146 CEMICS-2.0-CDR20                      
147 CEMICS-2.0-CDR8                       
148 CEMICS-Ref                            
149 PEP_1p5C_full_NDC                     
150 PEP_1p5C_full_eff                     
151 PEP_1p5C_full_goodpractice            
152 PEP_1p5C_full_netzero                 
153 PEP_1p5C_red_eff                      
154 PEP_2C_full_NDC                       
155 PEP_2C_full_eff                       
156 PEP_2C_full_goodpractice              
157 PEP_2C_full_netzero                   
158 PEP_2C_red_NDC                        
159 PEP_2C_red_eff                        
160 PEP_2C_red_goodpractice               
161 PEP_2C_red_netzero                    
162 SMP_1p5C_Def                          
163 SMP_1p5C_Sust                         
164 SMP_1p5C_early                        
165 SMP_1p5C_lifesty                      
166 SMP_1p5C_regul                        
167 SMP_2C_Def                            
168 SMP_2C_Sust                           
169 SMP_2C_early                          
170 SMP_2C_lifesty                        
171 SMP_2C_regul                          
172 SMP_REF_Def                           
173 SMP_REF_Sust                          
174 CEDS                                  
175 IEA Energy Statistics (r2017)         
176 Sky                                   
177 SSP4-19   
```

```{r eval = FALSE}

carbonTax <- seq(0,6,2) # ($ per tC) 
reduction <- seq(0,7,1)

my.fun <- function(reduction, carbonTax){

emission <- 10 - reduction
marginal <- seq(0,7, 1)
total <- cumsum(marginal)
tax <- emission * carbonTax
TotalPlusTax <- total + tax
return(TotalPlusTax)
}


data <- outer(reduction, carbonTax, my.fun)

my.xlab <- "Emissions reduced (tC)"
my.ylab <- "Costs of reductions + Tax (CAD)"

my.col <- rev(viridis::plasma(n = ncol(data)))

par(mar = c(5,5,1,1))
matplot(x = reduction, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("Carbon Tax =", carbonTax, "CAD/tC", sep = " ")
legend("bottomright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r eval = FALSE}
# Plant A
reduction <- seq(0,7, 1)
carbonTax <- 4 # ($ per tC) 
emission <- 10 - reduction
marginal <- seq(0,7, 1)
total <- cumsum(marginal)
tax <- emission * carbonTax
plantA <- total + tax 

# Plant B
reduction <- seq(0,7, 1)
carbonTax <- 4 # ($ per tC) 
emission <- 10 - reduction
marginal <- seq(0,7, 1) * 2
total <- cumsum(marginal)
tax <- emission * carbonTax
plantB <- total + tax 



par(mfrow = c(1,1), mar = c(5,5,1,1))
plot(reduction, plantA, type = "l", ylim = c(30,70),
     xlab = "Emissions reduced (tC)",
     ylab = "Costs of reductions + Tax (CAD)")

lines(reduction, plantB, col = "red")

my.col <- c("black", "red")

legend.text <- paste("Carbon Tax =", carbonTax, "CAD/tC", sep = " ")
legend("topleft", legend = c("Plant A", "Pant B (marginal costs 2 x that of Plant A)"), pch = 16, col = my.col, bty = "n")

```

